
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
```

Description:
- Outcome: hospitalizations with or without ADRD (sex-age standardized & sex-age-dual standardized)
- Ensure counties match in all objects
- Merge in geometry
- Remove counties with zero Black person-years (must do before creating adjacency matrix)
- Create county adjacency matrix

```{r}
########## user input ########## 

# outcome (adrd or nonadrd)
#outcome <- "adrd"
outcome <- "nonadrd"

# incorporate Medicaid eligibility into expected counts?
#dual <- TRUE # yes, standardize by age/sex/dual
dual <- FALSE # no, standardize by age/sex
```

```{r, message=FALSE, warning=FALSE}
## load packages ----
library(tidyverse)
library(magrittr)
library(sf)
library(ggthemes)
library(stringr)
```


## Shapefiles

```{r}
## read county shp files
county_sf <- read_sf("data/county/cb_2015_us_county/cb_2015_us_county_20m.shp") %>%
  filter(!STATEFP %in% c("02","15", "66", "72", "60", "69", "78")) %>% 
  mutate(county = as.numeric(GEOID)) %>%
  arrange(county)

#Oglala Lakota County, SD. Shannon County, SD (FIPS code = 46113) was renamed Oglala Lakota County and assigned anew FIPS code (46102) effective in 2014
county_sf$county[county_sf$county == 46102] <- 46113 

# get adjacent counties from shapefile
county_adj_list <-  st_touches(x = county_sf, y = county_sf, sparse = TRUE)

# this will become character. convert back to integer
names(county_adj_list) <- county_sf$county

# format as dataframe
county_adj_df <- do.call(rbind, lapply(names(county_adj_list), function(name) {
  data.frame(
    fips_i = rep(name, length(county_adj_list[[name]])),
    fips_j = names(county_adj_list)[county_adj_list[[name]]]
  )
}))

# change fips to numeric
county_adj_df$fips_i <- as.numeric(county_adj_df$fips_i)
county_adj_df$fips_j <- as.numeric(county_adj_df$fips_j)

# any counties without borders?
zero_borders <- names(which(lengths(county_adj_list) == 0))
zero_borders
```

```{r}
# inspect counties without borders
county_sf |> 
  filter(county %in% c(53055, 25019)) # two island counties

fips_sanjuan <- zero_borders[1] %>% as.numeric()
fips_nantucket <- zero_borders[2] %>% as.numeric()
```

```{r}
# manually add borders between these counties and the closest county

# San Juan, WA --> add border with Skagit county
fips_skagit <- county_sf |> filter(NAME == "Skagit") |>
  pull(county)

# Nantucket, MA --> add border with Dukes county
# get the fips for this county
fips_dukes <- county_sf |> filter(NAME == "Dukes") |>
  pull(county)

# make a new df with the new borders
# two rows for each border--one with county A in fips_i and county B in fips_j, one with county B in fips_i and county A in fips_j
new_borders <- data.frame(fips_i = c(fips_sanjuan, fips_skagit, 
                                     fips_nantucket, fips_dukes),
                          fips_j = c(fips_skagit, fips_sanjuan, 
                                     fips_dukes, fips_nantucket))

# bind these rows to adjacency
county_adj_df <- rbind(county_adj_df, new_borders)
```

```{r}
# add new columns to adjacency

# new county column
county_adj_df %<>%
  mutate(county = fips_i)

# df to link state and county
fips_to_state <- county_sf %>% 
  st_drop_geometry %>%
  select(county, STATEFP) %>%
  mutate(fips = county) %>%
  select(-county)

# new state columns
county_adj_df %<>%
  merge(fips_to_state, 
        by.x = "fips_i", by.y = "fips") %>%
  rename(state_fips_i = STATEFP) %>%
  merge(fips_to_state,
        by.x = "fips_j", by.y = "fips") %>%
  rename(state_fips_j = STATEFP)


dim(county_sf)
length(unique(county_adj_df$fips_i))
length(unique(county_adj_df$fips_j))
length(unique(county_adj_df$county))

counties_adj_ = unique(county_adj_df$county)
counties_shp_ = unique(county_sf$county)

setdiff(counties_shp_, counties_adj_)
setdiff(counties_adj_, counties_shp_)
```

## Outcome data

```{r}
# load outcome data
if(dual){
  county_df <- read_csv(paste0("data/symlinks/scratch/", outcome, "_rates_sensitivity.csv"))
} else {
  county_df <- read_csv(paste0("data/symlinks/scratch/", outcome, "_rates_main.csv"))
}

# filter out states
county_df <- county_df %>% 
    filter(!state %in% c(2, 15, 66, 72, 60, 69, 78)) %>%
    arrange(race, county)

counties_ = unique(county_df$county)

length(counties_) #3109
setdiff(counties_, counties_adj_) 
#when using 2010 crosswalk 51515 51560
#when using 2015 crosswalk 46102
setdiff(counties_adj_, counties_) 
#when using 2010 crosswalk 8014
#when using 2015 crosswalk none
```

```{r}
# Shannon County, SD (FIPS code = 46113) was renamed Oglala Lakota County and assigned a new FIPS code (46102) effective in 2014
zip_county <- read.csv("data/county/zip_county_2015.csv")

zip_county %>% 
  filter(county %in% c(46113, 46102))
```

```{r}
# change fips
county_df$county[county_df$county == 46102] <- 46113

# get # of enrollees and # of admissions by strata
if(dual){
  county_df %<>% 
    group_by(county, year, race, sex, age_grp, dual, state) %>% 
    summarise(n_enrollees = sum(n_enrollees), 
              n_admissions = sum(n_admissions)) %>% 
  ungroup()
} else {
  county_df %<>% 
    group_by(county, year, race, sex, age_grp, state) %>% 
    summarise(n_enrollees = sum(n_enrollees), 
              n_admissions = sum(n_admissions)) %>% 
  ungroup()
}
```

```{r}
counties_ = unique(county_df$county)

length(counties_)
setdiff(counties_, counties_adj_)
setdiff(counties_adj_, counties_)
```













```{r}
lapply(county_df, summary)
```

## Remove counties with zero Black person-years

```{r}
# remove counties with zero black person-years over the whole time period
# must do this *before* creating c_idx dummies and adjacency for stan

# get counties with zero person-years
# note: n_enrollees = NA means zero person-years
county_zero_b <- county_df %>%
  group_by(race, county) %>%
  filter(all(is.na(n_enrollees))) %>%
  distinct(county) %>%
  pull(county)

length(county_zero_b)

# save these counties so we can plot them later
#write_rds(county_zero_b, "data/intermediate/counties_zero_black_person_years.rds")
```

```{r}
# remove rows for these counties from data (both white and black rows)
county_df %<>%
  filter(!county %in% county_zero_b)

# and remove from county adjacency
county_adj_df %<>%
  filter(!fips_i %in% county_zero_b & !fips_j %in% county_zero_b)
length(unique(county_adj_df$county))

# and remove from county_sf, but save the geometry first
county_sf_zero_b <- county_sf %>%
  filter(county %in% county_zero_b)
write_rds(county_sf_zero_b, "data/symlinks/scratch/county_sf_zero_b.rds")

county_sf %<>%
  filter(!county %in% county_zero_b)
length(unique(county_sf$county))
```

# Save county-level data

```{r}
if(dual){
  write_rds(county_df, paste0("data/symlinks/scratch/", outcome, "-dual_county_df.rds"))
} else {
  write_rds(county_df, paste0("data/symlinks/scratch/", outcome, "_county_df.rds"))
}
```



## Adjacency matrix

Note: only need to run everything below this once! The county adjacency matrix is the same for both outcomes and for the main and sensitivity analyses.  


```{r}
####
# Create IDs and dummies
####

## Adding dummy FIPS ----
add_dummy_county_ids <- function(county_adj_df) {
  ## Just returns a County Adjacency matrix with two new columns -- each
  ## representing a dummy variable from 1:n_counties
  
  ## Sort by FIPS
  new_adj <- dplyr::arrange(county_adj_df, fips_i, fips_j)
  
  ## Make a dummy dictionary that with FIPS as the key and a dummy as value.
  dummies <- as.list(1:length(unique(new_adj$fips_i)))
  names(dummies) <- unique(new_adj$fips_i)
  
  ## Then make a dummy for node_i and node_j
  # need to keep as.character()!
  new_adj$dummy_i <-
    unlist(dummies[as.character(new_adj$fips_i)], use.names = FALSE)
  new_adj$dummy_j <-
    unlist(dummies[as.character(new_adj$fips_j)], use.names = FALSE)
  
  return(new_adj)
}

add_dummy_state_ids <- function(county_adj_df) {
  ## Just returns a County Adjacency matrix with two new columns -- each
  ## representing a dummy variable from 1:n_states
  
  ## Sort by FIPS
  new_adj <- dplyr::arrange(county_adj_df, fips_i, fips_j)
  
  ## Make a dummy dictionary that with FIPS as the key and a dummy as value.
  dummies <- as.list(1:length(unique(new_adj$state_fips_i)))
  names(dummies) <- unique(new_adj$state_fips_i)
  
  ## Then make a dummy for node_i and node_j
  new_adj$dummy_state_i <-
    unlist(dummies[new_adj$state_fips_i], use.names = FALSE)
  new_adj$dummy_state_j <-
    unlist(dummies[new_adj$state_fips_j], use.names = FALSE)
  
  return(new_adj)
}

county_adj_df <- add_dummy_county_ids(county_adj_df)
county_adj_df <- add_dummy_state_ids(county_adj_df)

county_adj_df
```

```{r}
####
# Obtain a table of county IDs
####

## Save a unique mapping of FIPS to state and county dummy codes ----
county_fips_df <- county_adj_df %>%
    dplyr::select(
        county = county, 
        fipschar = fips_i,
        fips_st = state_fips_i,
        c_idx = dummy_i,
        s_idx = dummy_state_i
    ) %>%
    dplyr::distinct()
county_fips_df
```

```{r}
length(unique(county_fips_df$county))
length(unique(county_df$county))
```

```{r}
## Save a sparse adjacency representation (for Stan) ----

make_county_adj_matrix <- function(county_adj_df) {
    ## Converts a County Adjacency dataframe into a traditional adjacency matrix
    ## of size n_counties x n_counties with 1 representing neighbors.
    
    ## Number of neighbors
    num <- county_adj_df %>%
        dplyr::group_by(dummy_i) %>%
        dplyr::summarize(neighbors_i = dplyr::n()) %>%
        dplyr::pull(neighbors_i)
    
    ## List of neighbors
    adj <- county_adj_df$dummy_j
    
    ## Create an appropriately sized matrix
    n_nodes <- length(num)
    A <- matrix(0L, nrow = n_nodes, ncol = n_nodes)
    
    ## Loop through num and adj and assign 1 to neighbors
    start_ix <- end_ix <- 0
    for (node in seq_along(num)) {
        for (n_neigh in num[node]) {
            start_ix  <- end_ix + 1
            end_ix    <- end_ix + n_neigh
            
            neighbors <- adj[start_ix:end_ix]
            
            A[node, neighbors] <- 1L
        }
    }
    return(A)
}

make_county_adj_sparse_list <- function(adj_matrix) {
  ## This is taken from mbjoseph CARstan repo -- returns a sparse
  ## representation of an adjacency matrix in the format that Stan needs
  ## Number of neighbors
  D_sparse <- rowSums(adj_matrix)
  
  ## Kyle Foreman script -- undirected graph so second line removes dupes
  W_sparse <- which(adj_matrix == 1, arr.ind = TRUE)
  W_sparse <- W_sparse[W_sparse[, 1] < W_sparse[, 2], ]
  
  ## Get eigenvalues of D^(-.5) * W * D^(-.5) for determinant computations
  invsqrtD <- diag(1 / sqrt(D_sparse))
  quadformDAD <- invsqrtD %*% adj_matrix %*% invsqrtD
  lambdas <- eigen(quadformDAD)$values
  
  ## Number of edges
  W_n <- nrow(W_sparse)
  
  return(list(
    D_sparse = D_sparse,
    W_sparse = W_sparse,
    lambdas = lambdas,
    W_n = W_n
  ))
}

county_adj_matrix <- make_county_adj_matrix(county_adj_df)
county_adj_sparse_list <- make_county_adj_sparse_list(county_adj_matrix)
names(county_adj_sparse_list)
```

```{r}
# Use return_sparse_parts(A) for these next ones
D_sparse = county_adj_sparse_list$D_sparse
# neighbors per node
W_sparse = county_adj_sparse_list$W_sparse
# adjacent pairs
lambda = county_adj_sparse_list$lambdas
# eigenvalues
W_n = nrow(county_adj_sparse_list$W_sparse)
           
length(D_sparse)
dim(W_sparse)
length(lambda)
W_n
```

```{r}
# confirm that county adjacency makes sense

# get the number of borders for each county
n_borders <- county_adj_df %>%
  group_by(county) %>%
  summarise(n_borders = n()) %>%
  mutate(n_borders = factor(n_borders, levels = c(seq(1, 20))))

# merge with geometry
check_adj <- merge(county_sf, n_borders, by = "county")

# plot number of borders
check_adj %>%
  ggplot() +
  geom_sf(aes(fill = n_borders), lwd = 0.1) +
  labs(fill = "Number of borders") +
  theme_map()
```

```{r}
# save results
write_rds(county_sf, "data/symlinks/scratch/county_sf.rds")
write_rds(county_adj_sparse_list, "data/symlinks/scratch/county_adj_sparse_list.rds")
write_rds(county_fips_df, "data/symlinks/scratch/county_fips_df.rds")
```
