
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
```

```{r, message=FALSE, warning=FALSE}
## Load libraries ----
library(tidyverse)
library(magrittr)
library(ggplot2)
library(sf)
#library(feather)
#library(cowplot)
library(RColorBrewer)
library(viridis)
library(ggthemes)
library(knitr)
library(kableExtra)
library(xtable)
```

```{r}
## Read data ----
adrd_county_df <- read_rds("data/symlinks/scratch/adrd_county_df_dual.rds")
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
county_fips_df <- read_rds("data/symlinks/scratch/county_fips_df.rds")

length(unique(adrd_county_df$county))
length(unique(county_sf$county))
length(unique(county_fips_df$county))

counties_shp_ = unique(county_sf$county)
counties_adrd_ = unique(adrd_county_df$county)
counties_ = unique(county_fips_df$county)

setdiff(counties_shp_, counties_adrd_)
setdiff(counties_adrd_, counties_shp_)
setdiff(counties_, counties_shp_)
setdiff(counties_shp_, counties_)
```

```{r}
## Subset ----
##  Subset to population of interest --
adrd_county_df %<>%
    filter(age_grp %in% c("[65,75)", "[75,85)", ">85"))

##  Replace NA with 0 --
adrd_county_df$n_enrollees[is.na(adrd_county_df$n_enrollees)] <- 0 
adrd_county_df$n_admissions[is.na(adrd_county_df$n_admissions)] <- 0 

sum(adrd_county_df$n_enrollees)
sum(adrd_county_df$n_admissions)

counties_shp_ = unique(county_sf$county)
counties_adrd_ = unique(adrd_county_df$county)

setdiff(counties_shp_, counties_adrd_)
setdiff(counties_adrd_, counties_shp_)
```


```{r}
# get basic stats from data (before rounding)

# number of hospitalizations with ADRD by race
adrd_county_df %>%
  group_by(race) %>%
  summarise(total_hosp = sum(n_admissions, na.rm = TRUE))
```

```{r}
# get sex and age distribution of person-years for each race

# table with % of person-years in each race that fall in each stratum
# sex_age_dist_tab <- adrd_county_df %>%
#   group_by(race, sex, age_grp) %>%
#   summarise(total_py = sum(n_enrollees)) %>%
#   ungroup() %>%
#   # group_by(race) %>%
#   # mutate(pct_py = round(100 * total_py / sum(total_py), 1)) %>%
#   spread(key = race, value = total_py) %>%
#   mutate(sex = ifelse(sex == 1, "Male", "Female"),
#          age_grp = ifelse(age_grp == "[65,75)", "65-74", age_grp),
#          age_grp = ifelse(age_grp == "[75,85)", "75-84", age_grp),
#          age_grp = ifelse(age_grp == ">85", "85+", age_grp),
#          white_pct = round(100 * `1` / sum(`1`), 1),
#          black_pct = round(100 * `2` / sum(`2`), 1)) %>%
#   # paste raw and % together
#   mutate(py_white = paste0(round(`1`, 0), " (", white_pct, " %)"),
#          py_black = paste0(round(`2`, 0), " (", black_pct, " %)")) %>%
#   select(-c(`1`, `2`, white_pct, black_pct)) %>%
#   arrange(age_grp) %>%
#   rename("Sex" = "sex",
#          "Age group" = "age_grp",
#          "White person-years" = "py_white",
#          "Black person-years" = "py_black")

# # table with % of person-years in white vs. black for each stratum 
# sex_age_dist_tab <- adrd_county_df %>%
#   group_by(race, sex, age_grp) %>%
#   summarise(total_py = sum(n_enrollees)) %>%
#   group_by(sex, age_grp) %>%
#   mutate(pct_py = round(100 * total_py / sum(total_py), 1)) %>%
#   mutate(total_pct_py = paste0(scales::comma(round(total_py, 0)), 
#                                " (", pct_py, "\\%)")) %>%
#   group_by(sex, age_grp) %>%
#   mutate(total_white_black_py = scales::comma(sum(total_py))) %>%
#   select(-c(total_py, pct_py)) %>%
#   spread(key = race, value = total_pct_py) %>%
#   ungroup() %>%
#   mutate(sex = ifelse(sex == 1, "Male", "Female"),
#          age_grp = ifelse(age_grp == "[65,75)", "aged 65-74", age_grp),
#          age_grp = ifelse(age_grp == "[75,85)", "aged 75-84", age_grp),
#          age_grp = ifelse(age_grp == ">85", "aged 85+", age_grp),
#          Stratum = paste0(sex, ", ", age_grp)) %>%
#   arrange(age_grp) %>%
#   rename("White" = `1`,
#          "Black" = `2`,
#          "Total" = total_white_black_py) %>%
#   select(c("Stratum", "White", "Black", "Total"))
# 
# # write_rds(sex_age_dist_tab, 
# #           file = paste0("results/tables/intermediate/sex_age_dist.rds"))
# # # table written to tex in script 10
# 
# # modify first column to include indent (easier for copy/paste)
# sex_age_dist_tab$Stratum <- paste0("\\hspace{20pt}", sex_age_dist_tab$Stratum)
# 
# # get table in latex
# # paste this table into the other table in overleaf
# print(xtable(sex_age_dist_tab, type = "latex"), 
#       file = "results/tables/final/sex_age_dist.tex", 
#       sanitize.text.function = identity,
#       include.rownames = FALSE)
```


# ```{r}
# ## Plot number of enrollees across years ----
# adrd_county_df %>% 
#   mutate(race = as.character(race)) %>% 
#   group_by(year, race) %>% 
#   summarise(n_enrollees = sum(n_enrollees)) %>% 
#   ggplot() + 
#   geom_line(aes(x = year, y = n_enrollees, color = race))
# ```
# 
# ```{r}
# ## Plot number of adrd admissions across years ----
# adrd_county_df %>% 
#   mutate(race = as.character(race)) %>% 
#   group_by(year, race) %>% 
#   summarise(n_admissions = sum(n_admissions)) %>% 
#   ggplot() + 
#   geom_line(aes(x = year, y = n_admissions, color = race))
# ```
# 
# ```{r}
# ## Number of enrollees by county-sex-age_grp split by race group ----
# adrd_county_df %>% 
#   mutate(race = as.character(race)) %>% 
#   ggplot() + 
#   geom_density(aes(x = n_enrollees, color = race))
# ```
# 
# ```{r}
# ## Number of enrollees by county-sex-age_grp split by race group ----
# adrd_county_df %>% 
#   filter(n_enrollees < 500) %>% 
#   mutate(race = as.character(race)) %>% 
#   ggplot() + 
#   geom_histogram(aes(x = n_enrollees, color = race)) +
#   facet_grid(~race)
# ```
# 
# ```{r}
# ## Adrd blind-to-counties rates for each race-sex-age_grp----
# adrd_county_df %>% 
#   group_by(race, sex, age_grp) %>% 
#   summarise(n_enrollees = sum(n_enrollees), 
#             n_admissions = sum(n_admissions),
#             adrd_rate = n_admissions / n_enrollees * 1e5) %>% 
#   mutate(race_sex = paste(race, sex, sep = "_")) %>% 
#   ggplot() +
#   geom_boxplot(aes(x = age_grp, y = adrd_rate)) + 
#   geom_jitter(aes(x = age_grp, y = adrd_rate, col = race_sex))
# ```
# 
# ```{r}
# ## Adrd county rates for each race-sex-age_grp ----
# adrd_county_df %>% 
#   group_by(county, race, sex, age_grp) %>% 
#   summarise(n_enrollees = sum(n_enrollees), 
#             n_admissions = sum(n_admissions),
#             adrd_rate = n_admissions / n_enrollees * 1e5) %>% 
#   mutate(race_sex = paste(race, sex, sep = "_")) %>% 
#   ggplot() +
#   geom_boxplot(aes(x = age_grp, y = adrd_rate)) + 
#   geom_jitter(aes(x = age_grp, y = adrd_rate, col = race_sex)) + 
#   facet_wrap(~race)
#```

```{r}
adrd_county_df %>% 
  group_by(race, sex, age_grp, dual) %>% 
  summarise(n_enrollees = sum(n_enrollees), 
            n_admissions = sum(n_admissions),
            adrd_rate = n_admissions / n_enrollees * 100) %>% 
  group_by(sex, age_grp, dual) %>% 
  summarise(adrd_rate = mean(adrd_rate))
```

```{r}
# xx <- adrd_county_df
# 
# xx$sex[adrd_county_df$sex == 1] = "male"
# xx$sex[adrd_county_df$sex == 2] = "female"
# 
# ## Reference adrd rates for each sex/age_grp blind to race ----
# xx %>% 
#   mutate(race = as.character(race)) %>% 
#   group_by(race, sex, age_grp) %>% 
#   summarise(n_enrollees = sum(n_enrollees), 
#             n_admissions = sum(n_admissions),
#             adrd_rate = n_admissions / n_enrollees * 100) %>% 
#   group_by(sex, age_grp) %>% 
#   summarise(adrd_rate = mean(adrd_rate)) %>%  
#   ggplot() + 
#   geom_point(aes(x = age_grp, y = adrd_rate, color = sex)) + 
#   labs(title = "ADRD reference rates", x = "Age group", y = "ADRD rate") + 
#   theme_bw()
# #ggsave("results/figures/adrd_reference_rates.png")
```

```{r}
# ## look at proportions by race
# prettyNum(prop.table(tapply(adrd_county_df$n_enrollees, adrd_county_df$race, sum)))
```

```{r}
# xx <- adrd_county_df
# 
# ## Observed county-level ADRD counts
# xx$race[xx$race == 1] = "white"
# xx$race[xx$race == 2] = "black"
# 
# xx %<>% 
#   group_by(county, race) %>% 
#   summarise(n_admissions = sum(n_admissions))
# 
# counties_shp_ = unique(county_sf$county)
# counties_xx_w_ = unique(xx$county[xx$race == "white"])
# counties_xx_b_ = unique(xx$county[xx$race == "black"])
# 
# setdiff(counties_shp_, counties_xx_w_)
# setdiff(counties_xx_w_, counties_shp_)
# setdiff(counties_shp_, counties_xx_b_)
# setdiff(counties_xx_b_, counties_shp_)
```

```{r}
# length(unique(xx$county))
# sum(is.na(xx$race))
```

```{r}
# yy <- county_sf %>% 
#   select(county) %>% 
#   left_join(xx) 
# 
# length(unique(yy$county))
# sum(is.na(yy$race))
# sum(is.na(yy$n_admissions))
```

```{r}
# yy %>%
#   filter(race == "white") %>% 
#   st_simplify() %>%
#   ggplot() +
#   geom_sf(aes(fill = n_admissions), alpha = 1, lwd = 0.0)
# 
# yy %>%
#   filter(race == "black") %>% 
#   st_simplify() %>%
#   ggplot() +
#   geom_sf(aes(fill = n_admissions), alpha = 1, lwd = 0.0)
```

```{r}
# n_admissions_ <- yy$n_admissions
# myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(10), 
#                            limits = c(min(n_admissions_, na.rm = TRUE), max(n_admissions_, na.rm = TRUE)))
# 
# yy  %>% 
#   ggplot() + 
#   geom_sf(aes(fill = n_admissions), alpha = 1, lwd = 0.0) +
#   facet_wrap(~race, ncol = 2) + 
#   labs(title = "Observed") + 
#   guides(fill = guide_legend(title = "ADRD counts")) + 
#   theme_classic() + 
#   theme(legend.position = "bottom") + 
#   sc
# #the yellow looks like white, try another visualization package
```

```{r}
# yy %>% 
#   ggplot() + 
#   geom_sf(aes(fill = n_admissions), alpha = 1, lwd = 0.0) +
#   facet_wrap(~race, ncol = 2) + 
#   labs(title = "Observed") + 
#   guides(fill = guide_legend(title = "ADRD counts")) + 
#   theme_classic() + 
#   scale_fill_viridis(name = "n_admissions", option = "D")
```

* adrd hospitalizations entangle prevalence and severity, specially in "small" counties 
* biased measurement error? noise systematically affects certain groups
* even if not affecting certain groups, measurement error decreases the power of the model
* variance of the inference, will not detect statistical differences in the output


## Observed vs expected

```{r}
# stage 1
# get race-specific national rates for each sex/age combination 
exp_stage1_df <- adrd_county_df %>%
    dplyr::group_by(race, sex, age_grp, dual) %>% # race, sex, age_grp
    dplyr::summarize(
        n_admissions = sum(n_admissions),
        n_enrollee_years = sum(n_enrollees),
        exp_rate_stage1 = n_admissions / n_enrollee_years
    )

# stage 2
# get overall national rates for each sex/age combination 
exp_stage2_df <- adrd_county_df %>%
    dplyr::group_by(sex, age_grp, dual) %>% # sex, age_grp
    dplyr::summarize(
        n_admissions = sum(n_admissions),
        n_enrollee_years = sum(n_enrollees),
        exp_rate_stage2 = n_admissions / n_enrollee_years
    )
```

```{r}
# make tables of standardized rates

#----- stage 1 -----#

exp_rates_stage1_tab <- exp_stage1_df %>%
  select(-c(n_enrollee_years, n_admissions)) %>%
  spread(key = "race", value = "exp_rate_stage1") %>%
  mutate(sex = ifelse(sex == 1, "Male", "Female"),
         age_grp = ifelse(age_grp == "[65,75)", "65-74", age_grp),
         age_grp = ifelse(age_grp == "[75,85)", "75-84", age_grp),
         age_grp = ifelse(age_grp == ">85", "85+", age_grp),
         `1` = format(round(`1`, 3), nsmall = 3),
         `2` = format(round(`2`, 3), nsmall = 3)) %>%
  arrange(age_grp) %>%
  rename("Sex" = "sex",
         "Age group" = "age_grp",
         "Rate (White)" = "1",
         "Rate (Black)" = "2")
exp_rates_stage1_tab


#----- stage 2 -----#

exp_rates_stage2_tab <- exp_stage2_df %>%
  select(-c(n_enrollee_years, n_admissions)) %>%
  mutate(sex = ifelse(sex == 1, "Male", "Female"),
         age_grp = ifelse(age_grp == "[65,75)", "65-74", age_grp),
         age_grp = ifelse(age_grp == "[75,85)", "75-84", age_grp),
         age_grp = ifelse(age_grp == ">85", "85+", age_grp),
         exp_rate_stage2 = format(round(exp_rate_stage2, 3), nsmall = 3)) %>%
  arrange(age_grp) %>%
  rename("Sex" = "sex",
         "Age group" = "age_grp",
         "Rate (overall)" = "exp_rate_stage2")
exp_rates_stage2_tab


# join tables together (not saving right now)
exp_rates_tab <- left_join(exp_rates_stage1_tab, exp_rates_stage2_tab)


# save tables
write_rds(exp_rates_stage1_tab, 
          file = paste0("results/tables/adrd_expected_rates_race-specific_dual.rds"))
write_rds(exp_rates_stage2_tab, 
          file = paste0("results/tables/adrd_expected_rates_overall_dual.rds"))
# tables joined and written to tex in script 10
```


```{r}
## Aggregate over years ----
## Just tally up the number of person years and hospitalizations over all years
obs_df <- adrd_county_df %>%
    dplyr::group_by(county, race, sex, age_grp, dual) %>% #county, race, sex, age_grp
    dplyr::summarize(n_enrollees = sum(n_enrollees),
                     n_admissions = sum(n_admissions), 
                     obs_rate = n_admissions / n_enrollees,
                     # HPI and PM2.5 are the same across all years
                     med_house_income = first(med_house_income),
                     house_price_income_ratio = first(house_price_income_ratio),
                     pm25 = mean(pm25))

## fill NAs
obs_df %>% 
  filter(n_enrollees == 0) 

obs_df$obs_rate[is.na(obs_df$obs_rate)] <- 0
```

```{r}
## Merge with exp_df ----
adrd_ratios_df <- obs_df %>%
    dplyr::left_join(exp_stage1_df %>%
                         dplyr::select(race, sex, age_grp, dual, exp_rate_stage1),
                     by = c("race", "sex", "age_grp", "dual")) %>%
    dplyr::left_join(exp_stage2_df %>%
                         dplyr::select(sex, age_grp, dual, exp_rate_stage2),
                     by = c("sex", "age_grp", "dual"))

## Now aggregate age ----

# here is where dual drops out!! along with age and sex

adrd_ratios_df %<>% 
  mutate(
    expected_stage1 = n_enrollees * exp_rate_stage1, 
    expected_stage2 = n_enrollees * exp_rate_stage2,
    observed = n_admissions
  ) %>% 
  dplyr::group_by(county, race) %>%
  dplyr::summarize(
    person_years = sum(n_enrollees),
    expected_stage1 = sum(expected_stage1), 
    expected_stage2 = sum(expected_stage2), 
    observed = sum(observed),
    # HPI and PM2.5 are the same across all years --> just take mean
    med_house_income = mean(med_house_income),
    house_price_income_ratio = mean(house_price_income_ratio),
    pm25 = mean(pm25)) %>%
  dplyr::ungroup()
```

```{r}
adrd_ratios_df %>% 
  mutate(
    rel_ratio = observed / expected_stage1,
    race = as.character(race)
  ) %>% 
  ggplot() + 
  geom_density(aes(x = rel_ratio, color = race))

adrd_ratios_df %>% 
  mutate(
    rel_ratio = observed / expected_stage2,
    race = as.character(race)
  ) %>% 
  ggplot() + 
  geom_density(aes(x = rel_ratio, color = race))
```

```{r}
adrd_ratios_df %>% 
  ggplot() + 
  geom_point(aes(x = observed, y = expected_stage1, color = as.factor(race)), alpha = 0.5) + 
  scale_x_sqrt() + 
  scale_y_sqrt()
```

```{r}
adrd_ratios_df %>% 
  mutate(
    log_diff = log(observed) - log(expected_stage1),
    race = as.character(race)
    ) %>% 
  ggplot() + 
  geom_density(aes(x = log_diff, color = race))

adrd_ratios_df %>% 
  mutate(
    log_diff = log(observed) - log(expected_stage2),
    race = as.character(race)
    ) %>% 
  ggplot() + 
  geom_density(aes(x = log_diff, color = race))
```

```{r}
xx <- adrd_ratios_df
xx$race[xx$race == 1] = "white"
xx$race[xx$race == 2] = "black"

xx %<>%
  mutate(
    log_diff = log(observed/expected_stage2),
    log_diff = if_else(observed == 0, as.numeric(NA), log_diff)
    )

counties_shp_ = unique(county_sf$county)
counties_xx_w_ = unique(xx$county[xx$race == "white"])
counties_xx_b_ = unique(xx$county[xx$race == "black"])

setdiff(counties_shp_, counties_xx_w_)
setdiff(counties_xx_w_, counties_shp_)
setdiff(counties_shp_, counties_xx_b_)
setdiff(counties_xx_b_, counties_shp_)
```

```{r}
length(unique(xx$county))
sum(is.na(xx$race))
```

```{r}
yy <- county_sf %>% 
  select(county) %>% 
  left_join(xx) 

length(unique(yy$county))
sum(is.na(yy$race))
sum(is.na(yy$log_diff))
sum(yy$person_years == 0)
sum(yy$observed == 0)
```

```{r}
yy %>% 
  mutate(person_years = log(person_years)) %>% 
  ggplot() + 
  geom_sf(aes(fill = person_years), alpha = 1, lwd = 0.0) +
  scale_fill_viridis(name = "person_years", option = "D") +
  facet_wrap(~race, ncol = 2) + 
  labs(title = "log person years") + 
  theme(legend.position = "bottom")
```

```{r}
## person years is zero
yy %>%
  st_simplify() %>%
  mutate(zero_persons = (person_years == 0)) %>%
  ggplot() + 
  geom_sf(aes(fill = zero_persons), alpha = 1, lwd = 0.0) + 
  facet_grid(~race)
```

```{r}
## observed is zero
yy %>%
  st_simplify() %>%
  mutate(zero_observed = (observed == 0)) %>%
  ggplot() + 
  geom_sf(aes(fill = zero_observed), alpha = 1, lwd = 0.0) + 
  facet_grid(~race)
```

```{r}
yy %>%
  filter(race == "white") %>% 
  st_simplify() %>%
  ggplot() +
  geom_sf(aes(fill = log_diff), alpha = 1, lwd = 0.0)

yy %>%
  filter(race == "black") %>% 
  st_simplify() %>%
  ggplot() +
  geom_sf(aes(fill = log_diff), alpha = 1, lwd = 0.0)
```

```{r}
yy %>% 
  ggplot() + 
  geom_sf(aes(fill = log_diff), alpha = 1, lwd = 0.0) +
  scale_fill_viridis(name = "log_diff", option = "D") +
  facet_wrap(~race, ncol = 2) + 
  labs(title = "log(observed) - log(expected)") + 
  theme(legend.position = "bottom")
```

```{r}
## Prep data ----
adrd_ratios_df %<>% 
  mutate(
    black = as.integer(if_else(race == 2, 1, 0)),
    # expected = as.integer(ceiling(expected)), # expected values don't need to be integers!
    observed = as.integer(round(observed)), # round normally (must be integers)
    # person_years = as.integer(round(person_years)), # doesn't need to be integers
    log_expected_stage1 = log(expected_stage1 + 0.00001), # to avoid issues with expected = 0
    log_expected_stage2 = log(expected_stage2 + 0.00001) # to avoid issues with expected = 0
  ) %>% 
  left_join(county_fips_df)

# sort by race and then county
adrd_ratios_df %<>%
  arrange(race, county)

sum(is.na(adrd_ratios_df$c_idx))
sum(is.na(adrd_ratios_df$s_idx))
sum(is.na(adrd_ratios_df$observed))
sum(is.na(adrd_ratios_df$expected_stage1))
sum(is.na(adrd_ratios_df$expected_stage2))
sum(is.na(adrd_ratios_df$person_years))
```


## Look at PM2.5 and house-price-to-income ratio

```{r}
# PM2.5 distribution by race (weighted by person-years)
adrd_ratios_df %>%
  mutate(race = case_when(
    race == 1 ~ "White",
    race == 2 ~ "Black"
  )) %>%
  ggplot(aes(pm25, weight = person_years, col = (race))) +
  geom_density() +
  labs(x = "PM2.5", y = "Density", col = "")

# HPI ratio distribution by race (weighted by person-years)
adrd_ratios_df %>%
mutate(race = case_when(
    race == 1 ~ "White",
    race == 2 ~ "Black"
  )) %>%
  ggplot(aes(house_price_income_ratio, weight = person_years, col = factor(race))) +
  geom_density(bw = 0.4) +
  labs(x = "House-price-income ratio", y = "Density", col = "")
```


```{r}
write_rds(adrd_ratios_df, "data/symlinks/scratch/adrd_ratios_df_dual.rds")
```
