
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
```

```{r, message=FALSE, warning=FALSE}
## Load libraries ----
library(tidyverse)
library(magrittr)
library(sf)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(ggpubr)
```

## Load data

```{r}
# choose which model to use
model <- "m1"

## Read model results
stanfit_samples <- read_rds(paste0("data/intermediate/adrd-", model, "_stanfit_samples.rds"))
stanfit_summary <- read_rds(paste0("data/intermediate/adrd-", model, "_stanfit_summary.rds"))

## Read data ----
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
pos_med <- read_rds(paste0("data/intermediate/adrd-", model, "_pos_med.rds"))

pos_med_sf <- county_sf %>% 
  select(county) %>% 
  right_join(pos_med)

dim(county_sf)
dim(pos_med_sf)
colnames(pos_med_sf)
```

```{r}
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
county_fips_df <- read_rds("data/symlinks/scratch/county_fips_df.rds")
state_sf <- read_sf("data/county/tl_2015_us_state/tl_2015_us_state.shp")
state_sf %<>% 
  left_join(
    county_sf %>% 
      left_join(county_fips_df) %>% 
      st_drop_geometry() %>% 
      distinct(STATEFP, s_idx, st_abbr)
  ) %>% 
  filter(!is.na(s_idx))
```

```{r}
# change projection to Albers Equal Area Conic (good for maps of US)
pos_med_sf <- st_transform(pos_med_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
state_sf <- st_transform(state_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))

```


```{r}
# read in Great Lakes shapefile
# https://www.naturalearthdata.com/downloads/110m-physical-vectors/110mlakes-reservoirs/
lakes_sf <- read_sf("data/county/ne_110m_lakes/ne_110m_lakes.shp") |>
  filter(name_alt == "Great Lakes")

# to exclude Great Lakes from maps, include the following layer:
lakes_gg <- geom_sf(data = lakes_sf, fill = "white", col = NA) 

# to draw state borders, include the following layer:
states_gg <- geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1)

# to exclude lakes and draw state borders, include the following layer:
lakes_states_gg <- list(
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1),
  geom_sf(data = lakes_sf, fill = "white", col = "black"))

# or this for white state borders:
lakes_states_light_gg <- list(
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1),
  geom_sf(data = lakes_sf, fill = "white", col = "white"))
```

```{r}
# race labels for plotting
race_labels <- c(
  `1` = "White",
  `2` = "Black"
)
```

```{r}
# # define color palettes
# 
# # viridis D (colorblind safe!!)
# viridis_D <- viridis_pal(option = "D")

```


## Pre-modelling plots

```{r}
# plot % of person-years contributed by black people in each county
pos_med_sf %>%
  group_by(county) %>%
  summarise(pct_blk = 100 * (person_years[race == 2] /
                               (person_years[race == 1] + person_years[race == 2]))) %>%
  ggplot() +
  geom_sf(aes(fill = pct_blk), lwd = 0) +
  lakes_states_light_gg +
  labs(fill = "% of person-years \ncontributed by Black \nindividuals") +
  scale_fill_viridis(option = "D") +
  theme_map() +
  theme(legend.title = element_text(size = 10))
```


```{r}
# map PM2.5, PIR, median household income

pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = pm25), alpha = 1, lwd = 0.0) +
  labs(fill = "PM2.5 (ug/m3)") +
  scale_fill_viridis(option = "D") +
  lakes_states_light_gg +
  theme(legend.position = "bottom") +
  theme_map()
  
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = log10(house_price_income_ratio)), alpha = 1, lwd = 0.0) +
  labs(fill = "log(PIR)") +
  scale_fill_viridis(option = "D") +
  lakes_states_light_gg +
  theme(legend.position = "bottom") +
  theme_map()

pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = med_house_income), alpha = 1, lwd = 0.0) +
  labs(fill = "Median household \nincome") +
  scale_fill_viridis(option = "D") +
  lakes_states_light_gg +
  theme(legend.position = "bottom") +
  theme_map()
```

### Observed vs. expected

```{r}
# map of observed counts
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = observed), lwd = 0.03) +
  scale_fill_viridis(option = "D") +
  lakes_states_light_gg +
  labs(fill = "Observed counts") +
  facet_grid(~race) +
  theme(legend.position = "bottom") +
  theme_map()

# map of expected counts
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = expected), lwd = 0.03) +
  lakes_states_light_gg +
  scale_fill_viridis(option = "D") +
  labs(fill = "Expected counts") +
  facet_grid(~race) +
  theme(legend.position = "bottom") +
  theme_map()
```

### Observed/expected counts ratio

Note: y is log(observed) - log(expected) or log(observed/expected), so exp(y) is observed/expected. y is analogous to the standardized mortality ratio (SMR) in Kiang et al.

```{r}
# look at distribution of these values (exp(y))
pos_med_sf |>
  ggplot() +
  geom_density(aes(exp(y)))
```


```{r}
# map of y (observed/expected)

# note: it's better to plot y than log(y)!
# if zero hosp are observed and 1 is expected, log(y) = -Inf, but y = 0
# if zero hosp are expected, then y and log(y) will both be undefined

# plotting in a slightly tricky/work-around way so outliers don't mess up the color scale
# note: legend shows up to 3, but max value for exp(y) is 11!
#pos_med_sf %>% 
ggplot() +
  #geom_sf(aes(fill = exp(y)), lwd = 0.00) +
  # plot normal range values with gradient
  geom_sf(data = subset(pos_med_sf, exp(y) <= 3), aes(fill = exp(y)), lwd = 0.00) +
  # plot high outliers in yellow (max of gradient)
  # geom_sf(data = subset(pos_med_sf, exp(y) > 3), fill = "#feef71", lwd = 0.00) +
  # plot high outliers in pale yellow
  geom_sf(data = subset(pos_med_sf, exp(y) > 3), fill = "#FFFCC9", lwd = 0.00) +
  # plot missing values in gray (must do manually here)
  geom_sf(data = subset(pos_med_sf, is.na(y)), fill = "gray50", lwd = 0.00) +
  lakes_states_light_gg +
  labs(fill = "Observed/Expected") +
  scale_fill_viridis(option = "D") + 
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
        )

# better idea if I want to do this:
# change all vals >3 to 3 (obv not in original data, make new df first)
# make another column that is either ">3" or NA
# make a border around each county colored by this new column
#   give outliers borders, otherwise no border
# honestly I think this is fine, just make a little box with the light yellow that says >3
```


```{r}
# similar map of y (observed/expected)
# but now highlight where observed/expected and lambda (from model) are very different
pos_med_sf %>%
  mutate(
    res = y - pos_med_y,
    res_is_out = if_else(abs(res) > 2/3, "different", "similar")
  ) %>%
  ggplot() +
  geom_sf(aes(fill = exp(y), col = as.factor(res_is_out)), lwd = 0.2) +
  scale_fill_viridis(option = "D") +
  labs(col = "y & lambda are:") +
  facet_grid(~race) +
  theme(legend.position = "bottom") +
  theme_map()
```


## Plot model results

### Relative/standardized risks (posterior medians of lambda = pos_med_y)
The "predictions" are a smoothed version of the observations.

```{r}
# plot lambdas
pos_med_sf %>% 
  # exclude counties with zero black person-years
  mutate(pos_med_y = ifelse(person_years == 0, NA, pos_med_y)) %>%
  ggplot() +
  geom_sf(aes(fill = exp(pos_med_y)), lwd = 0.00) +
  lakes_states_light_gg  +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  labs(fill = "Standardized risk (lambda)") +
  scale_fill_viridis(option = "D", na.value = "gray50") + 
  theme_map() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
```

### State effects

```{r}
# plot state intercepts (nu)

param = rownames(stanfit_summary$summary)
nu = tibble(
  param = param[grep("nu", param)],
  med = stanfit_summary$summary[grep("nu", param),"50%"],
  s_idx = 1:length(grep("nu", param))
)

state_sf %>% 
  left_join(nu) %>% 
  ggplot() + 
  geom_sf(aes(fill = med), col = "white") +
  lakes_gg +
  labs(fill = "State intercept (nu)") +
  theme_map()
```

### White-specific, Black-specific, and shared (unscaled) risks (three risk surfaces)

```{r}
summary(pos_med$shared_unscaled)
hist(pos_med$shared_unscaled)
```

```{r}
summary(pos_med$specific)
hist(pos_med$specific)
hist(exp(pos_med$specific))
```

```{r}
# find outliers for phi
pos_med_sf %<>% 
  mutate(
    shared_unscaled_inliers = if_else(shared_unscaled > -1 & shared_unscaled < 1, 
                                      shared_unscaled, as.numeric(NA)),
    shared_unscaled_outliers = if_else(shared_unscaled <= -1 | shared_unscaled >= 1, 
                                       shared_unscaled, as.numeric(NA)),
    shared_unscaled_outliers = if_else(shared_unscaled_outliers <= -1, -1, shared_unscaled_outliers),
    shared_unscaled_outliers = if_else(shared_unscaled_outliers >= 1, 1, shared_unscaled_outliers)
  ) 

pos_med_sf %>% 
  st_drop_geometry() %>% 
  select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# look at outliers
filter(pos_med_sf, !is.na(shared_unscaled_outliers)) %>% 
  st_drop_geometry() %>% 
  select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# look at counties with zero person-years
filter(pos_med_sf, person_years == 0) %>% 
  st_drop_geometry() %>% 
  select(county, race, person_years, shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# if zero person years, make "specific" and "shared" columns NA
pos_med_sf %<>% 
  mutate(
    specific = if_else(person_years == 0, as.numeric(NA), specific),
    shared_unscaled_inliers = if_else(person_years == 0, as.numeric(NA), shared_unscaled_inliers),
    shared_unscaled_outliers = if_else(person_years == 0, as.numeric(NA), shared_unscaled_outliers)
  )
```


```{r}
# plot three risk surfaces

# map 1: don't exponentiate, include non-significant counties

# not sure about this b/c I'm ignoring values <-1 and >1

# set color palette
myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            limits = c(-1, 1),
#                            na.value = "gray70")
sc <- scale_fill_gradient2(low = "#000080", mid = "white", high = "#ff6600",
                         midpoint = 0)

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  lakes_states_gg +
  theme_map() +
  labs(subtitle = "White-specific (psi1)",
       fill = "Median posterior estimate") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# shared unscaled (phi)
p2 <- pos_med_sf %>% 
  filter(race == 1) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared_unscaled_inliers), lwd = 0.0) + 
  lakes_states_gg +
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  theme_map() + 
  labs(subtitle = "Unscaled shared (phi)",
       fill = "Median posterior estimate") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# psi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  lakes_states_gg +
  theme_map() + 
  labs(subtitle = "Black-specific (psi2)",
       fill = "Median posterior estimate") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc
  
# p <- cowplot::plot_grid(
#   p1 + theme(legend.position = "none"), 
#   p2 + theme(legend.position = "none"), 
#   p3  + theme(legend.position = "none"),
#   nrow = 3)
# 
# cowplot::plot_grid(
#   p, 
#   cowplot::get_legend(p1), 
#   ncol = 2
# )

# alternative to cowplot
ggarrange(p1, p2, p3, 
          common.legend = TRUE, 
          nrow = 1,
          legend = "bottom")
```

```{r}
# map 2: exponentiate and only include non-significant counties

# new color scale (since I'm exponentiating)
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            na.value = "gray70")
sc <- scale_fill_gradient2(low = "#000080", mid = "white", high = "#ff6600",
                         midpoint = 1)

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = exp(specific)), lwd = 0.0) + 
  lakes_states_gg +
  labs(subtitle = "White-specific (exp(psi1))",
       fill = "Median posterior estimate") +
  theme_map() +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# shared unscaled (phi)
p2 <- pos_med_sf %>% 
  mutate(shared_unscaled_inliers = ifelse(shared_unscaled_sig, shared_unscaled_inliers, NA)) %>%
  filter(race == 1) %>% #shared_unscaled is non-race specific, so choose one
  ggplot() +
  geom_sf(aes(fill = exp(shared_unscaled_inliers)), lwd = 0.0) + 
  lakes_states_gg +
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  labs(subtitle = "Unscaled shared (exp(phi))",
       fill = "Median posterior estimate") +
  theme_map() + 
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# psi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = exp(specific)), lwd = 0.0) + 
  lakes_states_gg +
  labs(subtitle = "Black-specific (exp(psi2))",
       fill = "Median posterior estimate") +
  theme_map() + 
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# arrange plots
ggarrange(p1, p2, p3, 
          common.legend = TRUE, 
          nrow = 1,
          legend = "bottom")
```



### Now split shared risk into two scaled maps

```{r}
shared_ <- c(pos_med$shared, pos_med$specific)

myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = myPalette(100), 
                           limits = c(min(shared_), max(shared_)))

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.00) + 
  lakes_gg +
  theme_map() +
  labs(subtitle = "white specific") + 
  theme(legend.title = element_blank()) + 
  sc

# scaled phi for white people
p2 <- pos_med_sf %>% 
  filter(race == 1) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared), lwd = 0.00) + 
  lakes_gg +
  theme_map() + 
  labs(subtitle = "shared white") + 
  sc

# scaled phi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared), lwd = 0.00) + 
  lakes_gg +
  theme_map() + 
  labs(subtitle = "shared black") + 
  sc

# psi for black people
p4 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.00) + 
  lakes_gg +
  theme_map() + 
  labs(subtitle = "black specific") +
  sc
  
p <- cowplot::plot_grid(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3  + theme(legend.position = "none"),
  p4  + theme(legend.position = "none"),
  nrow = 2, 
  ncol = 2)

cowplot::plot_grid(
  p, 
  cowplot::get_legend(p1), 
  ncol = 2, 
  rel_widths = c(0.8, 0.2)
)
```


#############################################################
#############################################################
#############################################################

## New maps from Lauren!

```{r}
# recreate figure 3

# plot distribution of risk ratios
pos_med_sf %>%
  ggplot() +
  geom_density(aes(risk_ratio))
```

```{r}
# if zero person years, make "risk_ratio" column NA for BOTH white and black rows

# get counties with zero person-years for black people
county_zero_b <- pos_med_sf %>%
  filter(person_years == 0) %>%
  pull(county)

# make risk ratio NA for white and black rows
pos_med_sf %<>% 
  mutate(
    risk_ratio = if_else(county %in% county_zero_b, as.numeric(NA), risk_ratio)
  )
```

```{r}
# map of risk ratios

library(paletteer)
library(ggthemes)

# need to change color scale to show <1 and >1
# myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            limits = c(min(pos_med_sf$risk_ratio), max(pos_med_sf$risk_ratio)),
#                            na.value = "gray70")

# trying a discrete diverging color palette (but needs to be centered at 1!)
# new_sc <- scale_fill_gradientn(colors = rev(paletteer_c("ggthemes::Red-Blue-White Diverging", 20)[c(1:10, 14)]), 
#                            limits = c(min(pos_med_sf$risk_ratio), max(pos_med_sf$risk_ratio)),
#                            na.value = "gray70")

# trying a continuous diverging color palette 
library(colorspace)
sc2 <- scale_fill_continuous_diverging(palette = "Green", 
                                           na.value = "gray70",
                                           mid = 1#,
                                           #limits = c(0.7, 2.5),
                                           #breaks = c(seq(0.7, 1, by = 0.1), seq(1.1, 2.5, by = 0.2))
                                           )


sc3 <- scale_fill_gradient2(low = "#710193", mid = "white", high = "darkgreen",
                         midpoint = 1,
                         limits = c(0, 2.5), 
                         trans = "log")

sc4 <- ggplot2::scale_fill_brewer(NULL, palette = "PRGn", direction = -1)

pos_med_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA))
  ) %>%
  select(county, geometry, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio), lwd = 0.1) + 
  #lakes_gg +
  states_gg +
  geom_sf(data = lakes_sf, fill = "white", col = "black") +
  labs(fill = "Risk ratio") +
  theme_map() +
  sc2
```

```{r}
# map of risk ratios but don't remove non-significant counties

# need to change color scale to show <1 and >1
myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = myPalette(100), 
                           limits = c(min(pos_med_sf$risk_ratio), max(pos_med_sf$risk_ratio)),
                           na.value = "gray70")

pos_med_sf %>% 
  # mutate(
  #   risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA))
  # ) %>%
  select(county, geometry, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio), lwd = 0.00) + 
  lakes_gg +
  labs(fill = "Risk ratio") +
  theme_map() +
  sc
```


```{r}
# dichotomize risk ratio

# what % of counties are significantly <1? significantly >1? (slow!!)
pos_med_sf %>%
  #slice_sample(prop = 0.05) %>%
  filter(race == 2) %>% # same risk ratio for white and black rows
  # select(county, geometry, risk_ratio, risk_ratio_sig) %>%
  # distinct() %>%
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    risk_ratio = ifelse(risk_ratio < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  count(risk_ratio, name = "count")

pos_med_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, NA),
    risk_ratio = ifelse(risk_ratio < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  select(county, geometry, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = factor(risk_ratio)), lwd = 0.00) + 
  lakes_gg +
  labs(fill = "") +
  theme_map()
```

```{r}
# same thing, but keep all the counties that aren't significant
# seems like they are all the counties with higher risk for black
pos_med_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  select(county, geometry, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = factor(risk_ratio)), lwd = 0.00) + 
  lakes_gg +
  labs(fill = "") +
  theme_map()
```



```{r}
# try plotting risk ratio with discrete bins (instead of continuous or dichotomous)

# # make a new column for discrete bins
# pos_med_sf %<>% 
#   mutate(
#     risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
#     risk_ratio_bins = cut(risk_ratio, 
#                           breaks = c(0, 1, 1.5, 2, Inf), 
#                           labels = c("<1", "1-1.5", "1.5-2", ">2")
#                           )
#   )

# # get all sig. RR values
# rr_vals <- pos_med_sf |>
#   filter(race == 1) |>
#   filter(!is.na(risk_ratio)) |>
#   pull(risk_ratio)

# I don't like quantiles!! need a bin for <1
#rr_11_bins <- quantile(rr_vals, seq(0, 1, 0.1))


# make a new column for discrete bins
bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2.2, 3)

# get bin labels from bin breaks (get mean)
bin_df <- embed(bin_breaks, 2) |> as.data.frame()
bin_means <- rowMeans(bin_df) |>
  format(nsmall = 1)

# new column for risk ratio
pos_med_sf %<>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    risk_ratio_bins = cut(risk_ratio, 
                          breaks = bin_breaks,
                          labels = bin_means
                          ))
# choose hex codes
# brewer.pal(n = 11, name = "PiYG")
# cols <- c("#8E0152", "#C51B7D", "#DE77AE", "#F1B6DA", "#FDE0EF", "#F7F7F7",
#           "#E6F5D0", "#B8E186", "#7FBC41", "#4D9221", "#276419")

pos_med_sf %>% 
  select(county, geometry, risk_ratio_bins) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
  #lakes_gg +
  states_gg +
  geom_sf(data = lakes_sf, fill = "white", col = "black") +
  labs(fill = "Risk Ratio \n(Black/White)") +
  theme_map() +
  theme(legend.position = "right", 
        #legend.justification = "right", 
        #legend.direction = "vertical",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  scale_fill_manual(values = c("#07da63", 
                               "#7ffbb5",
                               "#FDE0EF", 
                               "#fbc8e3", 
                               "#DE77AE", 
                               "#C51B7D", 
                               "#8E0152", 
                               "#400023"),
                    na.value = "gray50",
                    breaks = levels(unique(pos_med_sf$risk_ratio_bins))#,
                    #na.translate = F      # makes NA white
                    ) #+
  #guides(fill = guide_legend(nrow = 1))
  
#scale_fill_manual(values = c("#492050FF", "#7C4586FF", "#E1C5E8FF", "#4D9221",
#                              "#B6D7B6FF", "#81B981FF", "#3F9640FF", "#226823FF", "#023903FF"))
    #values = myPalette(5), 
                    #values = rev(c("brown2", "darkorange", "gold", "darkgreen")),
                    #values = c("#88CFA4", "#FFFFBF", "#FFA500", "#BF0A30"),
                    #na.value = "gray50")


```


```{r}
# change plot to be more like Matt's with legend on bottom 

pos_med_sf %>% 
  select(county, geometry, risk_ratio_bins) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
  #lakes_gg +
  states_gg +
  geom_sf(data = lakes_sf, fill = "white", col = "black") +
  labs(fill = "Risk Ratio (Black/White)") +
  theme_map() +
  theme(legend.position = "bottom", 
        #legend.justification = "right", 
        #legend.direction = "vertical",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  scale_fill_manual(values = c("#07da63", 
                               "#7ffbb5",
                               "#FDE0EF", 
                               "#fbc8e3", 
                               #"#fab0d6", 
                               "#DE77AE", 
                               "#C51B7D", 
                               "#8E0152", 
                               "#400023"),
                    na.value = "gray50",
                    breaks = levels(unique(pos_med_sf$risk_ratio_bins))) +
  guides(fill = guide_legend(
    title.position = "top",        
    nrow = 1,
            label.position = "bottom",
            keywidth = grid::unit(3.5, "lines"),
            keyheight = grid::unit(.5, "lines"),
            label.hjust = 0))


```

this might help with these pink/green plots!!

scale_color_gradientn(colours = c("blue", "blue", "white", "red", "red"),
                        values = c(0, 0.485, 0.5, 0.515, 1))
