
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, message=FALSE, warning=FALSE}
## Load libraries ----
library(tidyverse)
library(magrittr)
library(sf)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(pander)
library(data.table)
```

## Load data

```{r}
# choose which model to use
#model <- "m1"
model <- "m2"
#model <- "m3"

## Read model results
stanfit_samples <- read_rds(paste0("data/intermediate/hosp-", model, "_stanfit_samples.rds"))
stanfit_summary <- read_rds(paste0("data/intermediate/hosp-", model, "_stanfit_summary.rds"))

## Read data ----
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
pos_med <- read_rds(paste0("data/intermediate/hosp-", model, "_pos_med.rds"))

pos_med_sf <- county_sf %>% 
  select(county) %>% 
  right_join(pos_med)

# dim(county_sf)
# dim(pos_med_sf)
colnames(pos_med_sf)
```

```{r}
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
county_fips_df <- read_rds("data/symlinks/scratch/county_fips_df.rds")
state_sf <- read_sf("data/county/cb_2015_us_state/cb_2015_us_state_20m.shp")
state_sf %<>% 
  left_join(
    county_sf %>% 
      left_join(county_fips_df) %>% 
      st_drop_geometry() %>% 
      distinct(STATEFP, s_idx)
  ) %>% 
  filter(!is.na(s_idx))
```

```{r}
# change projection to Albers Equal Area Conic (good for maps of US)
pos_med_sf <- st_transform(pos_med_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
state_sf <- st_transform(state_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
```

```{r}
# race labels for plotting
race_labels <- c(
  `1` = "White",
  `2` = "Black"
)
```

```{r}
# make a new column for rate of hospitalization (observed/person-years)
pos_med_sf %<>% 
  mutate(hosp_rate_obs = observed/person_years)

# and a new column for expected rate of hospitalization (expected/person-years)
pos_med_sf %<>% 
  mutate(hosp_rate_exp = expected/person_years)
```


## Look at counties with zero person-years and counties with zero expected hospitalizations

```{r}
# counties with zero person-years
filter(pos_med_sf, person_years == 0) %>% 
  st_drop_geometry() %>% 
  select(county, race, person_years, expected) %>%
  nrow()
# 71 counties with zero Black person-years

# get # of counties with zero expected hospitalizations
filter(pos_med_sf, expected == 0) %>% 
  st_drop_geometry() %>% 
  select(county, race, person_years, expected) %>% 
  nrow()
# the only counties with zero expected hospitalizations have zero person-years

```

```{r}
# if zero black person years, make lambda, specific, shared, and risk_ratio columns NA for BOTH white and black rows

# FOR NOW: keeping all counties

# # get counties with zero person-years for black people
# county_zero_b <- pos_med_sf %>%
#   filter(person_years == 0) %>%
#   pull(county)
# 
# # make NA for white and black rows
# pos_med_sf %<>% 
#   mutate(
#     pos_med_y = if_else(county %in% county_zero_b, as.numeric(NA), pos_med_y),
#     shared = if_else(county %in% county_zero_b, as.numeric(NA), shared),
#     shared_unscaled = if_else(county %in% county_zero_b, as.numeric(NA), shared_unscaled),
#     specific = if_else(county %in% county_zero_b, as.numeric(NA), specific),
#     risk_ratio = if_else(county %in% county_zero_b, as.numeric(NA), risk_ratio)
#   )
```


## A basic table

```{r}
# function to get summary statistics
summarize_column <- function(df, 
                             column_name, 
                             param_name, 
                             exp_transform = FALSE,
                             group_race = TRUE) {
  result <- df %>%
    st_drop_geometry() %>% 
    group_by(race) %>%
    summarize(
      min = ifelse(exp_transform,
                   min(exp({{ column_name }}), na.rm = TRUE),
                   min({{ column_name }}, na.rm = TRUE)),
      p250 = ifelse(exp_transform,
                    stats::quantile(exp({{ column_name }}), .250, na.rm = TRUE),
                    stats::quantile({{ column_name }}, .250, na.rm = TRUE)),
      p500 = ifelse(exp_transform,
                    stats::median(exp({{ column_name }}), na.rm = TRUE),
                    stats::median({{ column_name }}, na.rm = TRUE)),
      p750 = ifelse(exp_transform,
                    stats::quantile(exp({{ column_name }}), .750, na.rm = TRUE),
                    stats::quantile({{ column_name }}, .750, na.rm = TRUE)),
      max = ifelse(exp_transform,
                   max(exp({{ column_name }}), na.rm = TRUE),
                   max({{ column_name }}, na.rm = TRUE))
    ) %>%
    mutate(param = param_name,
           min = round(min, 2),
           p250 = round(p250, 2),
           p500 = round(p500, 2),
           p750 = round(p750, 2),
           max = round(max, 2))
  return(result)
}

# apply function to get summary stats table
# note: I don't think the first two can come off of FASSE because they show small counts
summary_stats <- rbind(
  # summarize_column(pos_med_sf, 
  #                  column_name = person_years,
  #                  param_name = "person years"),
  # summarize_column(pos_med_sf, 
  #                  column_name = observed,
  #                  param_name = "observed"),
  summarize_column(df = pos_med_sf, 
                   column_name = y, 
                   param_name = "Observed/expected", 
                   exp_transform = TRUE),
  summarize_column(df = pos_med_sf, 
                   column_name = pos_med_y, 
                   param_name = "Standardized risk (lambda)", 
                   exp_transform = TRUE),
  summarize_column(df = pos_med_sf, 
                   column_name = risk_ratio, 
                   param_name = "Risk ratio",
                   group_race = FALSE)
  )

# make risk ratio a single row (same for both races)
summary_stats %<>%
  filter(!(param == "Risk ratio" & race == 2)) %>%
  mutate(race = ifelse(param == "Risk ratio", "Both", race))

# change race values to white and black
summary_stats %<>%
  mutate(race = ifelse(race == 1, "White", race)) %>%
  mutate(race = ifelse(race == 2, "Black", race)) %>%
  select(c(param, everything()))

# display without scientific notation
summary_stats[] <- lapply(summary_stats, function(x) if(is.numeric(x)) format(x, scientific = FALSE) else x)


# rename columns
summary_stats <- summary_stats |>
  rename(Parameter = param,
            Race = race,
            Minimum = min,
            `25th percentile` = p250,
            `50th percentile` = p500,
            `75th percentile` = p750,
            Maximum = max)
summary_stats |> pander()

# save table
write_rds(summary_stats, file = paste0("results/tables/summary_stats_", model, ".rds"))
```


### Alternatively, plot densities of these distributions

```{r}
# excluding lambda because that comes after modeling
# maybe compare all the lambdas across models separately

# I want to facet by parameter, so I need to gather the columns of interest
param_dist <- pos_med_sf |>
  st_drop_geometry() |>
  mutate(exp_y = exp(y),
         exp_pos_med_y = exp(pos_med_y)) |>
  gather(key = "parameter", value = "value", 
         c(person_years, 
           observed, 
           expected,
           hosp_rate_obs,
           hosp_rate_exp,
           exp_y)) |>
  select(county, race, parameter, value)

# look at NA values
# NA values are for ADRD rate the 71 
# param_dist |>
#   group_by(parameter) |>
#   summarise(num_NA = sum(is.na(value))) %>%
#   nrow()
# these are just the 71 counties with 0 black person-years, so I can ignore them

# I want to plot on log scale, so change any 0s to 0.000001
param_dist %<>%
  mutate(value = ifelse(value == 0, 0.000001, value))

# set order of plots
param_dist$parameter <- factor(param_dist$parameter,
                               levels = c("person_years", "observed", "expected", "exp_y", 
                                          "hosp_rate_obs", "hosp_rate_exp"))

# plot
param_dist |>
  filter(!is.na(value)) |>
  mutate(race = factor(ifelse(race == 1, "White", "Black"),
                       levels = c("White", "Black"))) %>% 
  ggplot() +
  geom_density(aes(value, col = as.factor(race))) +
  labs(x = "",
       y = "Density",
       col = "") +
  scale_x_log10(labels = scales::comma) +
  facet_wrap(~parameter, scales = "free", 
             labeller = as_labeller(c(hosp_rate_obs = "Observed ADRD rate",
                                      hosp_rate_exp = "Expected ADRD rate",
                                      #exp_pos_med_y = "Lambda",
                                      exp_y = "Observed hosp./expected hosp.",
                                      observed = "Observed hosp.",
                                      expected = "Expected hosp.",
                                      person_years = "Person years")
  )) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom")

# plotting x-axis on log scale!!

```


## Empirical variances

```{r}
readRDS(paste0("results/tables/empirical_variance_", model, ".rds")) |> pander()
```



## Understand the data/covariates

```{r}
# # plot % of person-years contributed by black people in each county
# pos_med_sf %>%
#   group_by(county) %>%
#   summarise(pct_blk = 100 * (person_years[race == 2] /
#                                (person_years[race == 1] + person_years[race == 2]))) %>%
#   ggplot() +
#   geom_sf(aes(fill = pct_blk), lwd = 0) +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   labs(fill = "% of person-years \ncontributed by Black \nindividuals") +
#   scale_fill_viridis(option = "D") +
#   theme_map() +
#   theme(legend.title = element_text(size = 10))
```


```{r}
# map PM2.5, PIR, median household income

# pos_med_sf %>%
#   ggplot() +
#   geom_sf(aes(fill = pm25), alpha = 1, lwd = 0.0) +
#   labs(fill = expression('PM'[2.5]*' (\u00B5g/m'^3*')')) +
#   scale_fill_viridis(option = "D") +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   theme(legend.position = "bottom") +
#   theme_map()
#   
# pos_med_sf %>%
#   ggplot() +
#   geom_sf(aes(fill = log10(house_price_income_ratio)), alpha = 1, lwd = 0.0) +
#   labs(fill = "log(house price-to- \nincome ratio)") +
#   scale_fill_viridis(option = "D") +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   theme(legend.position = "bottom") +
#   theme_map()
# 
# pos_med_sf %>%
#   ggplot() +
#   geom_sf(aes(fill = med_house_income), alpha = 1, lwd = 0.0) +
#   labs(fill = "Median household \nincome") +
#   scale_fill_viridis(option = "D") +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   theme(legend.position = "bottom") +
#   theme_map()
```


## Pre-modeling visualizations of observed and expected

### Observed rate of hospitalization  (observed/person-years)

```{r}
# try capping at 50%--otherwise we're just seeing really small counties
# alternatively, could restrict this to counties with a certain number of person-years observed
pos_med_sf %>%
  mutate(hosp_rate_obs = ifelse(hosp_rate_obs > 0.5, 0.5, hosp_rate_obs)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_obs * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Rate of hosp. (%)") +
  scale_fill_viridis(option = "D") +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()

# could also try lumping together white and black (this just tells us where there are more ADRD hospitalizations)
```


### Expected rate of hospitalization  (expected/person-years)

```{r}
# try capping at 50%--otherwise we're just seeing really small counties
# alternatively, could restrict this to counties with a certain number of person-years
pos_med_sf %>%
  mutate(hosp_rate_exp = ifelse(hosp_rate_exp > 0.6, 0.6, hosp_rate_exp)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_exp * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Expected rate of \nhosp.  (%)") +
  scale_fill_viridis(option = "D") +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()

# could also try lumping together white and black (this just tells us where there are more ADRD hospitalizations)
```

### Observed vs. expected

```{r}
# map of observed counts
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = observed), lwd = 0.03) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = "Observed counts") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()

# map of expected counts
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = expected), lwd = 0.03) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  labs(fill = "Expected counts") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()
```

### Observed/expected counts ratio

Note: y is log(observed) - log(expected) or log(observed/expected), so exp(y) is observed/expected. exp(y) is analogous to the standardized mortality ratio (SMR) in Kiang et al.


```{r}
# map of observed/expected = exp(y)

# note: it's better to plot y than log(y)!
# if zero hosp are observed and 1 is expected, log(y) = -Inf, but y = 0
# (if zero hosp are expected, then y and log(y) will both be undefined)

# plotting in a slightly tricky/work-around way so outliers don't mess up the color scale
# note: legend shows up to 3, but max value for exp(y) is 11!
#pos_med_sf %>% 
ggplot() +
  #geom_sf(aes(fill = exp(y)), lwd = 0.00) +
  # plot normal range values with gradient
  geom_sf(data = subset(pos_med_sf, exp(y) <= 3), aes(fill = exp(y)), lwd = 0.00) +
  # plot high outliers in yellow (max of gradient)
  geom_sf(data = subset(pos_med_sf, exp(y) > 3), fill = "#EDFF00", lwd = 0.00) +
  # plot high outliers in pale yellow
  # geom_sf(data = subset(pos_med_sf, exp(y) > 3), fill = "#FFFCC9", lwd = 0.00) +
  # plot missing values in gray (must do manually here)
  geom_sf(data = subset(pos_med_sf, is.na(y)), fill = "gray60", lwd = 0.00) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = "Observed/expected") +
  scale_fill_viridis(option = "D") + 
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
        )

# note: the only counties in gray here have 0 black person-years, meaning that they also have zero expected hospitalizations , meaning observed/expected is undefined


# better idea if I want to do this:
# change all vals >3 to 3 (obv not in original data, make new df first)
# make another column that is either ">3" or NA
# make a border around each county colored by this new column
#   give outliers borders, otherwise no border
# honestly I think this is fine, just make a little box with the light yellow that says >3
```



## Model results

### Relative/standardized risks (posterior medians of lambda = pos_med_y)
The "predictions" are a smoothed version of the observations.

```{r}
# plot lambdas
pos_med_sf %>% 
  # # don't plot lambda for zero black person-years for black counties only
  # mutate(pos_med_y = ifelse(person_years == 0, NA, pos_med_y)) %>%
  ggplot() +
  geom_sf(aes(fill = exp(pos_med_y)), lwd = 0.00) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  labs(fill = expression('Smoothed standardized risk ('*lambda*')')) +
  scale_fill_viridis(option = "D", na.value = "gray60") + 
  theme_map() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
```

```{r}
# look at distribution of lambda by race
pos_med_sf |>
  mutate(race = ifelse(race == 1, "White", "Black")) %>% 
  ggplot() +
  geom_density(aes(x = exp(pos_med_y), color = race), linewidth = 1) + 
  labs(x = "Spatially smoothed standardized risk (lambda)", 
       y = "Density",
       col = "") +
  theme_classic()
```

```{r}
# now compare raw ratios (observed/expected--unsmoothed) to model results (lambda--smoothed)
pos_med_sf %>%
  mutate(
    res = y - pos_med_y,
    res_is_out = if_else(abs(res) > 2/3, "different", "similar")
  ) %>%
  ggplot() +
  geom_sf(aes(#fill = exp(y), 
              fill = as.factor(res_is_out)), lwd = 0.1) +
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Raw ratio & \nlambda are:") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()
# makes sense that they are more different for Black individuals (more to smooth)
```


### State effects

```{r}
# plot state intercepts (eta)
# note: eta was referred to as nu in the original code!

param <- rownames(stanfit_summary$summary)
nu <- tibble(
  param = param[grep("nu", param)],
  med = stanfit_summary$summary[grep("nu", param),"50%"],
  s_idx = 1:length(grep("nu", param))
)

state_sf %>% 
  left_join(nu) %>% 
  ggplot() + 
  geom_sf(aes(fill = med), col = "white") +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = expression('State intercept ('*eta*')')) +
  theme_map()
```

### White-specific, Black-specific, and shared (unscaled) risks (three risk surfaces)

```{r}
# summary(pos_med$shared_unscaled)
# hist(pos_med$shared_unscaled)
```

```{r}
# summary(pos_med$specific)
# hist(pos_med$specific)
# hist(exp(pos_med$specific))
```

```{r}
# # find outliers for phi
# pos_med_sf %<>% 
#   mutate(
#     shared_unscaled_inliers = if_else(shared_unscaled > -1 & shared_unscaled < 1, 
#                                       shared_unscaled, as.numeric(NA)),
#     shared_unscaled_outliers = if_else(shared_unscaled <= -1 | shared_unscaled >= 1, 
#                                        shared_unscaled, as.numeric(NA)),
#     shared_unscaled_outliers = if_else(shared_unscaled_outliers <= -1, -1, shared_unscaled_outliers),
#     shared_unscaled_outliers = if_else(shared_unscaled_outliers >= 1, 1, shared_unscaled_outliers)
#   ) 

# pos_med_sf %>% 
#   st_drop_geometry() %>% 
#   select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# # look at outliers
# filter(pos_med_sf, !is.na(shared_unscaled_outliers)) %>% 
#   st_drop_geometry() %>% 
#   select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers) %>%
#   nrow()
```


```{r}
# plot three risk surfaces
# first, include non-significant counties

# set color palette
# not sure about old palette b/c it ignores values <-1 and >1 and isn't colorblind safe
# myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            limits = c(-1, 1),
#                            na.value = "gray70")
sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0)

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() +
  labs(subtitle = expression('White-specific ('*psi[1]*')'),
       fill = "Median posterior estimate") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# shared unscaled (phi)
p2 <- pos_med_sf %>% 
  filter(race == 1) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared_unscaled), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  theme_map() + 
  labs(subtitle = expression('Unscaled shared (\u03A6)'),
       fill = "Median posterior estimate") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# psi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(subtitle = expression('Black-specific ('*psi[2]*')'),
       fill = "Median posterior estimate") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc
  
# p <- cowplot::plot_grid(
#   p1 + theme(legend.position = "none"),
#   p2 + theme(legend.position = "none"),
#   p3  + theme(legend.position = "none"),
#   nrow = 3)
# 
# cowplot::plot_grid(
#   p,
#   cowplot::get_legend(p1),
#   ncol = 2
# )

# alternative to cowplot
ggarrange(p1, p2, p3,
          common.legend = TRUE,
          nrow = 1,
          legend = "bottom")
```

### Now exclude non-significant counties

```{r}
# new color scale (since I'm exponentiating)
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            na.value = "gray70")
# sc <- scale_fill_gradient2(low = "#000080", mid = "white", high = "#ff6600", na.value = "gray40",
#                          midpoint = 0)
sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0)

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(subtitle = expression('White-specific ('*psi[1]*')'),
       fill = "Median posterior estimate") +
  theme_map() +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# shared unscaled (phi)
p2 <- pos_med_sf %>% 
  mutate(shared_unscaled = ifelse(shared_unscaled_sig, shared_unscaled, NA)) %>%
  filter(race == 1) %>% #shared_unscaled is non-race specific, so choose one
  ggplot() +
  geom_sf(aes(fill = shared_unscaled), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  labs(subtitle = expression('Unscaled shared (\u03A6)'),
       fill = "Median posterior estimate") +
  theme_map() + 
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# psi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(subtitle = expression('Black-specific ('*psi[2]*')'),
       fill = "Median posterior estimate") +
  theme_map() + 
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  sc

# arrange plots
ggarrange(p1, p2, p3, 
          common.legend = TRUE, 
          nrow = 1,
          legend = "bottom")
```



### Now split shared risk into two scaled maps

```{r}
# split phi map into phi*delta and phi/delta

# sc <- scale_fill_gradient2(low = "#000080", mid = "white", high = "#ff6600", na.value = "gray40",
#                          midpoint = 0)
sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0)

# # vector of all shared and specific values
# shared_ <- c(pos_med$shared, pos_med$specific)
# myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            limits = c(min(shared_), max(shared_)))

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() +
  labs(subtitle = expression('White-specific ('*psi[1]*')'),
       fill = "Median posterior \nestimate") +
  theme(plot.subtitle = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) + 
  sc

# scaled phi for white people
p2 <- pos_med_sf %>% 
  filter(race == 1) %>%
  mutate(shared = ifelse(shared_sig, shared, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = shared), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(subtitle = expression('Scaled shared for White (\u03A6*\u03B4)'),
       fill = "Median posterior \nestimate") +
  theme(plot.subtitle = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) + 
  sc

# scaled phi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>%
  mutate(shared = ifelse(shared_sig, shared, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = shared), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(subtitle = expression('Scaled shared for White (\u03A6/\u03B4)'),
       fill = "Median posterior \nestimate") +
  theme(plot.subtitle = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) + 
  sc

# psi for black people
p4 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(subtitle = expression('Black-specific ('*psi[2]*')'),
       fill = "Median posterior \nestimate") +
  theme(plot.subtitle = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) + 
  sc
  
p <- cowplot::plot_grid(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3  + theme(legend.position = "none"),
  p4  + theme(legend.position = "none"),
  nrow = 2, 
  ncol = 2)

cowplot::plot_grid(
  p, 
  cowplot::get_legend(p1), 
  ncol = 2, 
  rel_widths = c(0.8, 0.2)
)
```


### Risk ratios

```{r}
# plot distribution of risk ratios
pos_med_sf %>%
  ggplot() +
  geom_histogram(aes(risk_ratio), col = "white") +
  geom_vline(xintercept = 1, lty = 2) +
  labs(x = "Risk ratio", y = "Count") +
  theme_bw()
```

```{r}
# dichotomize risk ratio--where is white vs. black risk higher?

# make a table--what % of counties are significantly <1? significantly >1?
RR_tab <- pos_med_sf %>%
  #slice_sample(prop = 0.05) %>%
  filter(race == 2) %>% # same risk ratio for white and black rows
  # select(county, geometry, risk_ratio, risk_ratio_sig) %>%
  # distinct() %>%
  select(-geometry) %>%
  as.data.frame() %>%
  mutate(
    `Risk Ratio` = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    `Risk Ratio` = ifelse(`Risk Ratio` < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  count(`Risk Ratio`, name = "Count")

# print table
RR_tab |> pander()

```

```{r}
# plot dichotomized risk ratio
pos_med_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, NA),
    risk_ratio = ifelse(risk_ratio < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  select(county, geometry, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = factor(risk_ratio)), lwd = 0.06) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "") +
  scale_fill_manual(values = c("Risk ratio < 1" = "#2171b5", 
                               "Risk ratio > 1" = "#fdae6b"),
                    na.value = "gray90") +
  theme_map()
```


```{r}
# some attempts at continuous color scales (challenging!)

# library(paletteer)
# library(ggthemes)

# # this works, but the blue doesn't get dark enough
# sc <- scale_fill_gradientn(colors = rev(paletteer_c("ggthemes::Red-Blue-White Diverging", 
#                                                     20)[c(1:10, 14)]),
#                            limits = c(min(pos_med_sf$risk_ratio), max(pos_med_sf$risk_ratio)),
#                            na.value = "gray70")

# # this is blue to brown for some reason
# library(colorspace)
# sc <- colorspace::scale_fill_continuous_diverging(palette = "Green", 
#                                            na.value = "gray70",
#                                            mid = 1#,
#                                            #limits = c(0.7, 2.5),
#                                            #breaks = c(seq(0.7, 1, by = 0.1), seq(1.1, 2.5, by = 0.2))
#                                            )

# # this works, but again, purple doesn't get dark enough
# sc <- scale_fill_gradient2(low = "#710193", mid = "white", high = "darkgreen",
#                          midpoint = 1)

# Error: Continuous value supplied to discrete scale
# sc <- ggplot2::scale_fill_brewer(NULL, palette = "PRGn", direction = -1)

# # map of risk ratios
# pos_med_sf %>% 
#   mutate(
#     risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA))
#   ) %>%
#   select(county, geometry, risk_ratio) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio), lwd = 0.1) + 
#   lakes_states_gg +
#   labs(fill = "Risk ratio") +
#   theme_map() +
#   sc
```


```{r}
# try plotting risk ratio with discrete bins (instead of continuous or dichotomous)

# set boundaries for each discrete bin
# may need to change!! should be consistent across models and include all RR
# bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2.2, 3)
bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, 2.2, 2.4, 3.6)


# get bin labels from bin breaks (get mean)
bin_df <- embed(bin_breaks, 2) |> as.data.frame()
bin_means <- rowMeans(bin_df) |> format(nsmall = 1)

# new column for risk ratio
pos_med_sf %<>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    risk_ratio_bins = cut(risk_ratio, 
                          breaks = bin_breaks,
                          labels = bin_means
                          ))
# choose hex codes manually from PiYG palette
# brewer.pal(n = 11, name = "PiYG")
# cols <- c("#8E0152", "#C51B7D", "#DE77AE", "#F1B6DA", "#FDE0EF", "#F7F7F7",
#           "#E6F5D0", "#B8E186", "#7FBC41", "#4D9221", "#276419")

# plot
pos_med_sf %>% 
  select(county, geometry, risk_ratio_bins) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Risk Ratio \n(Black/White)") +
  theme_map() +
  theme(legend.position = "right", 
        #legend.justification = "right", 
        #legend.direction = "vertical",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  scale_fill_manual(
    # values = c("#07da63",
    #            "#7ffbb5",
    #            "#FDE0EF",
    #            "#fbc8e3",
    #            "#DE77AE",
    #            "#C51B7D",
    #            "#8E0152",
    #            "#400023"),
    # values = c("#08306b",
    #            "#2171b5",
    #            "#feedde",
    #            "#fdd0a2", 
    #            "#fdae6b", 
    #            "#fd8d3c", 
    #            "#e6550d", 
    #            "#a63603"),
    values = c("#08306b",
               "#2171b5",
               "#fee6ce",
               "#fdd0a2", 
               "#fdae6b", 
               "#fd8d3c", 
               "#f16913",
               "#d94801",
               "#a63603", 
               "#7f2704"),
    na.value = "gray80",
    breaks = levels(unique(pos_med_sf$risk_ratio_bins))#,
    #na.translate = F      # makes NA white
  ) #+
  #guides(fill = guide_legend(nrow = 1))

# # or change legend to be along the bottom like Matt's
# pos_med_sf %>% 
#   select(county, geometry, risk_ratio_bins) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
#   #lakes_gg +
#   states_gg +
#   geom_sf(data = lakes_sf, fill = "white", col = "black") +
#   labs(fill = "Risk Ratio (Black/White)") +
#   theme_map() +
#   theme(legend.position = "bottom", 
#         #legend.justification = "right", 
#         #legend.direction = "vertical",
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   scale_fill_manual(values = c("#07da63", 
#                                "#7ffbb5",
#                                "#FDE0EF", 
#                                "#fbc8e3", 
#                                #"#fab0d6", 
#                                "#DE77AE", 
#                                "#C51B7D", 
#                                "#8E0152", 
#                                "#400023"),
#                     na.value = "gray50",
#                     breaks = levels(unique(pos_med_sf$risk_ratio_bins))) +
#   guides(fill = guide_legend(
#     title.position = "top",        
#     nrow = 1,
#             label.position = "bottom",
#             keywidth = grid::unit(3.5, "lines"),
#             keyheight = grid::unit(.5, "lines"),
#             label.hjust = 0))

```


