
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, message=FALSE, warning=FALSE}
## Load libraries ----
library(tidyverse)
library(magrittr)
library(sf)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(pander)
library(data.table)
```

## Load data

```{r}
# load list with all model results and bind into df
pos_med <- read_rds(paste0("data/intermediate/hosp_pos_med.rds"))
pos_med <- bind_rows(pos_med, .id = "model")
```

```{r}
## Read data ----
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
#pos_med <- read_rds(paste0("data/intermediate/hosp-", model, "_pos_med.rds"))


pos_med_sf <- county_sf %>% 
  select(county) %>% 
  right_join(pos_med)

# dim(county_sf)
# dim(pos_med_sf)
colnames(pos_med_sf)
```

```{r}
county_fips_df <- read_rds("data/symlinks/scratch/county_fips_df.rds")
state_sf <- read_sf("data/county/cb_2015_us_state/cb_2015_us_state_20m.shp")
state_sf %<>% 
  left_join(
    county_sf %>% 
      left_join(county_fips_df) %>% 
      st_drop_geometry() %>% 
      distinct(STATEFP, s_idx)
  ) %>% 
  filter(!is.na(s_idx))
```

```{r}
# change projection to Albers Equal Area Conic (good for maps of US)
pos_med_sf <- st_transform(pos_med_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
state_sf <- st_transform(state_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
```

```{r}
# race labels for plotting
race_labels <- c(
  `1` = "White",
  `2` = "Black"
)

# model labels for plotting
model_labels <- c(
  m2 = "Unadjusted",
  m2 = "Adjusted for\n price-to-income ratio",
  m3 = "Adjusted for PM\u2082.\u2085"
)

# race and model labels for plotting
race_model_labels <- c(
  `1` = "White",
  `2` = "Black",
  m2 = "Unadjusted",
  m2 = "Adjusted for\n price-to-income ratio",
  m3 = "Adjusted for PM\u2082.\u2085"
)
```

```{r}
# make a new column for rate of hospitalization (observed/person-years)
pos_med_sf %<>% 
  mutate(hosp_rate_obs = observed/person_years)

# and a new column for expected rate of hospitalization (expected/person-years)
pos_med_sf %<>% 
  mutate(hosp_rate_exp = expected/person_years)
```


## Pre-modeling visualizations of observed and expected

### Observed rate of hospitalization (observed/person-years)

```{r}
# try capping at 50%--otherwise we're just seeing really small counties
# alternatively, could restrict this to counties with a certain number of person-years observed
pos_med_sf %>%
  filter(model == "m2") %>%
  mutate(hosp_rate_obs = ifelse(hosp_rate_obs > 0.4, 0.4, hosp_rate_obs)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_obs * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Rate of hosp. (%)") +
  scale_fill_viridis(option = "D") +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()

# could also try lumping together white and black (this just tells us where there are more hosp hospitalizations)
```


### Expected rate of hospitalization with hosp (expected/person-years)

```{r}
# try capping at 50%--otherwise we're just seeing really small counties
# alternatively, could restrict this to counties with a certain number of person-years
pos_med_sf %>%
  filter(model == "m2") %>%
  mutate(hosp_rate_exp = ifelse(hosp_rate_exp > 0.4, 0.4, hosp_rate_exp)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_exp * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Expected rate of \nhosp. (%)") +
  scale_fill_viridis(option = "D") +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()

# could also try lumping together white and black (this just tells us where there are more hosp hospitalizations)
```

### Observed vs. expected

```{r}
# map of observed counts
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = observed), lwd = 0.03) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = "Observed counts") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()

# map of expected counts
pos_med_sf %>%
  ggplot() +
  geom_sf(aes(fill = expected), lwd = 0.03) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  labs(fill = "Expected counts") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme(legend.position = "bottom") +
  theme_map()
```

### Observed/expected counts ratio

Note: y is log(observed) - log(expected) or log(observed/expected), so exp(y) is observed/expected. exp(y) is analogous to the standardized mortality ratio (SMR) in Kiang et al.


```{r}
# map of observed/expected = exp(y)

# note: it's better to plot y than log(y)!
# if zero hosp are observed and 1 is expected, log(y) = -Inf, but y = 0
# (if zero hosp are expected, then y and log(y) will both be undefined)

pos_med_sf %>% 
  mutate(y = ifelse(exp(y) > 3, log(3), y)) %>%
  ggplot() +
  geom_sf(aes(fill = exp(y)), lwd = 0.00) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = "Observed/expected") +
  scale_fill_viridis(option = "A") + 
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
        )

# note: the only counties in gray here have 0 black person-years, meaning that they also have zero expected hospitalizations with hosp, meaning observed/expected is undefined
```



## Model results

### Relative/standardized risks (posterior medians of lambda = pos_med_y)
The "predictions" are a smoothed version of the observations.

```{r}
# plot lambdas
pos_med_sf %>% 
  # # don't plot lambda for zero black person-years for black counties only
  # mutate(pos_med_y = ifelse(person_years == 0, NA, pos_med_y)) %>%
  ggplot() +
  geom_sf(aes(fill = exp(pos_med_y)), lwd = 0.0) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(race ~ model,
             labeller = as_labeller(race_model_labels)
             ) +
  labs(fill = expression('Smoothed standardized risk ('*lambda*')')) +
  scale_fill_viridis(option = "A", 
                     na.value = "gray60") + 
  theme_map() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
```

```{r}
# now compare raw ratios (observed/expected--unsmoothed) to model results (lambda--smoothed)
pos_med_sf %>%
  mutate(
    res = y - pos_med_y,
    res_is_out = if_else(abs(res) > 2/3, "different", "similar")
  ) %>%
  ggplot() +
  geom_sf(aes(#fill = exp(y), 
              fill = as.factor(res_is_out)), lwd = 0.1) +
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Raw ratio & \nlambda are:") +
  facet_grid(race ~ model,
             labeller = as_labeller(race_model_labels)
             ) +
  theme(legend.position = "bottom") +
  theme_map()
# makes sense that they are more different for Black individuals (more to smooth)
```


### State effects

```{r}
# plot state intercepts (eta)
# note: eta was referred to as nu in the original code!

# param = rownames(stanfit_summary$summary)
# nu = tibble(
#   param = param[grep("nu", param)],
#   med = stanfit_summary$summary[grep("nu", param),"50%"],
#   s_idx = 1:length(grep("nu", param))
# )
# 
# state_sf %>% 
#   left_join(nu) %>% 
#   ggplot() + 
#   geom_sf(aes(fill = med), col = "white") +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   labs(fill = expression('State intercept ('*eta*')')) +
#   theme_map()
```

### White-specific, Black-specific, and shared (unscaled) risks (three risk surfaces)

```{r}
# summary(pos_med$shared_unscaled)
# hist(pos_med$shared_unscaled)
```

```{r}
# summary(pos_med$specific)
# hist(pos_med$specific)
# hist(exp(pos_med$specific))
```

```{r}
# # find outliers for phi
# pos_med_sf %<>% 
#   mutate(
#     shared_unscaled_inliers = if_else(shared_unscaled > -1 & shared_unscaled < 1, 
#                                       shared_unscaled, as.numeric(NA)),
#     shared_unscaled_outliers = if_else(shared_unscaled <= -1 | shared_unscaled >= 1, 
#                                        shared_unscaled, as.numeric(NA)),
#     shared_unscaled_outliers = if_else(shared_unscaled_outliers <= -1, -1, shared_unscaled_outliers),
#     shared_unscaled_outliers = if_else(shared_unscaled_outliers >= 1, 1, shared_unscaled_outliers)
#   ) 

# pos_med_sf %>% 
#   st_drop_geometry() %>% 
#   select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# # look at outliers
# filter(pos_med_sf, !is.na(shared_unscaled_outliers)) %>% 
#   st_drop_geometry() %>% 
#   select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers) %>%
#   nrow()
```


```{r}
# plot three risk surfaces
# first, include non-significant counties

# set color palette
# not sure about old palette b/c it ignores values <-1 and >1 and isn't colorblind safe
# myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            limits = c(-1, 1),
#                            na.value = "gray70")
sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0)

# # psi for white people
# p1 <- pos_med_sf %>% 
#   filter(race == 1) %>% 
#   ggplot() +
#   geom_sf(aes(fill = specific), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   theme_map() +
#   labs(subtitle = expression('White-specific ('*psi[1]*')'),
#        fill = "Median posterior estimate") +
#   facet_wrap(~ model,
#              labeller = as_labeller(model_labels),
#              nrow = 3
#              ) +
#   theme(legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   sc
# 
# # shared unscaled (phi)
# p2 <- pos_med_sf %>% 
#   filter(race == 1) %>% #shared_unscaled is non-race specific
#   ggplot() +
#   geom_sf(aes(fill = shared_unscaled), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
#   # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
#   theme_map() + 
#   labs(subtitle = expression('Unscaled shared (\u03A6)'),
#        fill = "Median posterior estimate") +
#   facet_wrap(~ model,
#              labeller = as_labeller(model_labels),
#              nrow = 3
#              ) +
#   theme(legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   sc
# 
# # psi for black people
# p3 <- pos_med_sf %>% 
#   filter(race == 2) %>% 
#   ggplot() +
#   geom_sf(aes(fill = specific), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   theme_map() + 
#   labs(subtitle = expression('Black-specific ('*psi[2]*')'),
#        fill = "Median posterior estimate") +
#   facet_wrap(~ model,
#              labeller = as_labeller(model_labels),
#              nrow = 3
#              ) +
#   theme(legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   sc
#   
# # p <- cowplot::plot_grid(
# #   p1 + theme(legend.position = "none"),
# #   p2 + theme(legend.position = "none"),
# #   p3  + theme(legend.position = "none"),
# #   nrow = 3)
# # 
# # cowplot::plot_grid(
# #   p,
# #   cowplot::get_legend(p1),
# #   ncol = 2
# # )
# 
# # alternative to cowplot
# ggarrange(p1, p2, p3,
#           common.legend = TRUE,
#           nrow = 1,
#           legend = "bottom")

###

shared_unscaled_df <- pos_med_sf |>
  select(county, geometry, model, race, 
         specific, specific_sig,
         shared_unscaled, shared_unscaled_sig) |>
  gather(key = "spec_vs_shared", value = "estimate", specific, shared_unscaled) %>%
  # new sig column for estimate
  mutate(estimate_sig = ifelse(spec_vs_shared == "specific", 
                               specific_sig, shared_unscaled_sig)) |>
  # remove other sig columns for clarity
  select(-c(specific_sig, shared_unscaled_sig)) |>
  # change race for shared rows
  mutate(race = ifelse(spec_vs_shared == "shared_unscaled", "both", race)) %>%
  distinct() |>
  mutate(parameter = paste0(race, "_", spec_vs_shared)) 

# set levels for plotting
shared_unscaled_df$parameter <- factor(shared_unscaled_df$parameter,
                                       levels = c("1_specific", "both_shared_unscaled", 
                                                  "2_specific"))

# labels for facets
shared_unscaled_labels <- c(
  m2 = "Unadjusted",
  m2 = "Adjusted for\n price-to-income ratio",
  m3 = "Adjusted for PM\u2082.\u2085",
  "1_specific" = "White-specific\n (\u03C8\u2081)",
  "both_shared_unscaled" = "Unscaled shared\n (\u03A6)",
  "2_specific" = "Black-specific\n (\u03C8\u2082)"
)

# plot
shared_unscaled_df |>
  ggplot() +
  geom_sf(aes(fill = estimate), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(fill = "Median posterior estimate") +
  facet_grid(parameter ~ model,
             labeller = as_labeller(shared_unscaled_labels)
             ) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        strip.text = element_text(size = 10)) +
  sc
```

### Now exclude non-significant counties

```{r}
# new color scale (since I'm exponentiating)
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            na.value = "gray70")
# sc <- scale_fill_gradient2(low = "#000080", mid = "white", high = "#ff6600", na.value = "gray40",
#                          midpoint = 0)
sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0)

# # psi for white people
# p1 <- pos_med_sf %>% 
#   filter(race == 1) %>% 
#   mutate(specific = ifelse(specific_sig, specific, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = specific), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   labs(subtitle = expression('White-specific ('*psi[1]*')'),
#        fill = "Median posterior estimate") +
#   facet_wrap(~ model,
#              labeller = as_labeller(model_labels),
#              nrow = 3
#              ) +
#   theme_map() +
#   theme(legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   sc
# 
# # shared unscaled (phi)
# p2 <- pos_med_sf %>% 
#   mutate(shared_unscaled = ifelse(shared_unscaled_sig, shared_unscaled, NA)) %>%
#   filter(race == 1) %>% #shared_unscaled is non-race specific, so choose one
#   ggplot() +
#   geom_sf(aes(fill = shared_unscaled), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
#   # geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
#   labs(subtitle = expression('Unscaled shared (\u03A6)'),
#        fill = "Median posterior estimate") +
#   facet_wrap(~ model,
#              labeller = as_labeller(model_labels),
#              nrow = 3
#              ) +
#   theme_map() + 
#   theme(legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   sc
# 
# # psi for black people
# p3 <- pos_med_sf %>% 
#   filter(race == 2) %>% 
#   mutate(specific = ifelse(specific_sig, specific, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = specific), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   labs(subtitle = expression('Black-specific ('*psi[2]*')'),
#        fill = "Median posterior estimate") +
#   facet_wrap(~ model,
#              labeller = as_labeller(model_labels),
#              nrow = 3
#              ) +
#   theme_map() + 
#   theme(legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   sc
# 
# # arrange plots
# ggarrange(p1, p2, p3, 
#           common.legend = TRUE, 
#           nrow = 1,
#           legend = "bottom")


# plot
shared_unscaled_df |>
  mutate(estimate = ifelse(estimate_sig, estimate, NA)) |>
  ggplot() +
  geom_sf(aes(fill = estimate), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(fill = "Median posterior estimate") +
  facet_grid(parameter ~ model,
             labeller = as_labeller(shared_unscaled_labels)
             ) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        strip.text = element_text(size = 10)) +
  sc
```



### Now split shared risk into two scaled maps

```{r}
# split phi map into phi*delta and phi/delta

# sc <- scale_fill_gradient2(low = "#000080", mid = "white", high = "#ff6600", na.value = "gray40",
#                          midpoint = 0)
sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0)

# # vector of all shared and specific values
# shared_ <- c(pos_med$shared, pos_med$specific)
# myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
# sc <- scale_fill_gradientn(colors = myPalette(100), 
#                            limits = c(min(shared_), max(shared_)))

# # psi for white people
# p1 <- pos_med_sf %>% 
#   filter(race == 1) %>% 
#   mutate(specific = ifelse(specific_sig, specific, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = specific), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   theme_map() +
#   labs(subtitle = expression('White-specific ('*psi[1]*')'),
#        fill = "Median posterior \nestimate") +
#   theme(plot.subtitle = element_text(size = 10),
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) + 
#   sc
# 
# # scaled phi for white people
# p2 <- pos_med_sf %>% 
#   filter(race == 1) %>%
#   mutate(shared = ifelse(shared_sig, shared, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = shared), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   theme_map() + 
#   labs(subtitle = expression('Scaled shared for White (\u03A6*\u03B4)'),
#        fill = "Median posterior \nestimate") +
#   theme(plot.subtitle = element_text(size = 10),
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) + 
#   sc
# 
# # scaled phi for black people
# p3 <- pos_med_sf %>% 
#   filter(race == 2) %>%
#   mutate(shared = ifelse(shared_sig, shared, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = shared), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   theme_map() + 
#   labs(subtitle = expression('Scaled shared for Black (\u03A6/\u03B4)'),
#        fill = "Median posterior \nestimate") +
#   theme(plot.subtitle = element_text(size = 10),
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) + 
#   sc
# 
# # psi for black people
# p4 <- pos_med_sf %>% 
#   filter(race == 2) %>% 
#   mutate(specific = ifelse(specific_sig, specific, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = specific), lwd = 0.03) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   theme_map() + 
#   labs(subtitle = expression('Black-specific ('*psi[2]*')'),
#        fill = "Median posterior \nestimate") +
#   theme(plot.subtitle = element_text(size = 10),
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) + 
#   sc
#   
# p <- cowplot::plot_grid(
#   p1 + theme(legend.position = "none"), 
#   p2 + theme(legend.position = "none"), 
#   p3  + theme(legend.position = "none"),
#   p4  + theme(legend.position = "none"),
#   nrow = 2, 
#   ncol = 2)
# 
# cowplot::plot_grid(
#   p, 
#   cowplot::get_legend(p1), 
#   ncol = 2, 
#   rel_widths = c(0.8, 0.2)
# )


shared_df <- pos_med_sf |>
  select(county, geometry, model, race, 
         specific, specific_sig,
         shared, shared_sig) |>
  gather(key = "spec_vs_shared", value = "estimate", specific, shared) %>%
  # new sig column for estimate
  mutate(estimate_sig = ifelse(spec_vs_shared == "specific", 
                               specific_sig, shared_sig)) |>
  # remove other sig columns for clarity
  select(-c(specific_sig, shared_sig)) |>
  mutate(parameter = paste0(race, "_", spec_vs_shared)) 

# set levels for plotting
shared_df$parameter <- factor(shared_df$parameter,
                              levels = c("1_specific", "1_shared",
                                         "2_shared", "2_specific"))

# labels for facets
shared_labels <- c(
  m2 = "Unadjusted",
  m2 = "Adjusted for PIR",
  m3 = "Adjusted for PM2.5",
  "1_specific" = "White-specific\n (\u03C8\u2081)",
  "1_shared" = "Scaled shared for\n White (\u03A6*\u03B4)",
  "2_shared" = "Scaled shared for\n Black (\u03A6/\u03B4)",
  "2_specific" = "Black-specific\n (\u03C8\u2082)"
)

# plot
shared_df |>
  mutate(estimate = ifelse(estimate_sig, estimate, NA)) |>
  ggplot() +
  geom_sf(aes(fill = estimate), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  theme_map() + 
  labs(fill = "Median posterior estimate") +
  facet_grid(parameter ~ model,
             labeller = as_labeller(shared_labels)
             ) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        strip.text = element_text(size = 7)) +
  sc
```


### Risk ratios

```{r}
# dichotomize risk ratio--where is white vs. black risk higher?

# make a table--what % of counties are significantly <1? significantly >1?
RR_tab <- pos_med_sf %>%
  #slice_sample(prop = 0.05) %>%
  filter(race == 2) %>% # same risk ratio for white and black rows
  # select(county, geometry, risk_ratio, risk_ratio_sig) %>%
  # distinct() %>%
  select(-geometry) %>%
  as.data.frame() %>%
  mutate(
    `Risk Ratio` = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    `Risk Ratio` = ifelse(`Risk Ratio` < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  group_by(model) %>%
  count(`Risk Ratio`, name = "Count")

# print table
RR_tab |> pander()

```

```{r}
# plot dichotomized risk ratio
pos_med_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, NA),
    risk_ratio = ifelse(risk_ratio < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  select(county, geometry, model, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = factor(risk_ratio)), lwd = 0.06) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "") +
  facet_grid(~ model,
             labeller = as_labeller(model_labels)
             ) +
  scale_fill_manual(values = c("Risk ratio < 1" = "#2171b5", 
                               "Risk ratio > 1" = "#fdae6b"),
                    na.value = "gray90") +
  theme_map()
```


```{r}
# some attempts at continuous color scales (challenging!)

# library(paletteer)
# library(ggthemes)

# # this works, but the blue doesn't get dark enough
# sc <- scale_fill_gradientn(colors = rev(paletteer_c("ggthemes::Red-Blue-White Diverging", 
#                                                     20)[c(1:10, 14)]),
#                            limits = c(min(pos_med_sf$risk_ratio), max(pos_med_sf$risk_ratio)),
#                            na.value = "gray70")

# # this is blue to brown for some reason
# library(colorspace)
# sc <- colorspace::scale_fill_continuous_diverging(palette = "Green", 
#                                            na.value = "gray70",
#                                            mid = 1#,
#                                            #limits = c(0.7, 2.5),
#                                            #breaks = c(seq(0.7, 1, by = 0.1), seq(1.1, 2.5, by = 0.2))
#                                            )

# # this works, but again, purple doesn't get dark enough
# sc <- scale_fill_gradient2(low = "#710193", mid = "white", high = "darkgreen",
#                          midpoint = 1)

# Error: Continuous value supplied to discrete scale
# sc <- ggplot2::scale_fill_brewer(NULL, palette = "PRGn", direction = -1)

# # map of risk ratios
# pos_med_sf %>% 
#   mutate(
#     risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA))
#   ) %>%
#   select(county, geometry, risk_ratio) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio), lwd = 0.1) + 
#   lakes_states_gg +
#   labs(fill = "Risk ratio") +
#   theme_map() +
#   sc
```


```{r}
# try plotting risk ratio with discrete bins (instead of continuous or dichotomous)

# set boundaries for each discrete bin
# may need to change!! should be consistent across models and include all RR
# bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2.2, 3)
bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, 2.2, 2.4, 3.6)


# get bin labels from bin breaks (get mean)
bin_df <- embed(bin_breaks, 2) |> as.data.frame()
bin_means <- rowMeans(bin_df) |> format(nsmall = 1)

# new column for risk ratio
pos_med_sf %<>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    risk_ratio_bins = cut(risk_ratio, 
                          breaks = bin_breaks,
                          labels = bin_means
                          ))
# choose hex codes manually from PiYG palette
# brewer.pal(n = 11, name = "PiYG")
# cols <- c("#8E0152", "#C51B7D", "#DE77AE", "#F1B6DA", "#FDE0EF", "#F7F7F7",
#           "#E6F5D0", "#B8E186", "#7FBC41", "#4D9221", "#276419")

# plot
pos_med_sf %>% 
  select(county, geometry, model, risk_ratio_bins) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Risk Ratio \n(Black/White)") +
  facet_grid(~ model,
             labeller = as_labeller(model_labels)
             ) +
  theme_map() +
  theme(legend.position = "bottom", 
        #legend.justification = "right", 
        #legend.direction = "vertical",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  scale_fill_manual(
    # values = c("#07da63",
    #            "#7ffbb5",
    #            "#FDE0EF",
    #            "#fbc8e3",
    #            "#DE77AE",
    #            "#C51B7D",
    #            "#8E0152",
    #            "#400023"),
    # values = c("#08306b",
    #            "#2171b5",
    #            "#feedde",
    #            "#fdd0a2", 
    #            "#fdae6b", 
    #            "#fd8d3c", 
    #            "#e6550d", 
    #            "#a63603"),
    values = c("#08306b",
               "#2171b5",
               "#fee6ce",
               "#fdd0a2", 
               "#fdae6b", 
               "#fd8d3c", 
               "#f16913",
               "#d94801",
               "#a63603", 
               "#7f2704"),
    na.value = "gray80",
    breaks = levels(unique(pos_med_sf$risk_ratio_bins))#,
    #na.translate = F      # makes NA white
  ) +
  guides(fill = guide_legend(
    title.position = "top",
    nrow = 1,
            label.position = "bottom",
            keywidth = grid::unit(2.5, "lines"),
            keyheight = grid::unit(.5, "lines"),
            label.hjust = 0))

# # or change legend to be along the bottom like Matt's
# pos_med_sf %>% 
#   select(county, geometry, risk_ratio_bins) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
#   #lakes_gg +
#   states_gg +
#   geom_sf(data = lakes_sf, fill = "white", col = "black") +
#   labs(fill = "Risk Ratio (Black/White)") +
#   theme_map() +
#   theme(legend.position = "bottom", 
#         #legend.justification = "right", 
#         #legend.direction = "vertical",
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8)) +
#   scale_fill_manual(values = c("#07da63", 
#                                "#7ffbb5",
#                                "#FDE0EF", 
#                                "#fbc8e3", 
#                                #"#fab0d6", 
#                                "#DE77AE", 
#                                "#C51B7D", 
#                                "#8E0152", 
#                                "#400023"),
#                     na.value = "gray50",
#                     breaks = levels(unique(pos_med_sf$risk_ratio_bins))) +
#   guides(fill = guide_legend(
#     title.position = "top",        
#     nrow = 1,
#             label.position = "bottom",
#             keywidth = grid::unit(3.5, "lines"),
#             keyheight = grid::unit(.5, "lines"),
#             label.hjust = 0))

```


## Look at distributions of post-modeling variables across counties

### In a table

```{r}
# function to get summary statistics
summarize_column <- function(df, 
                             column_name, 
                             param_name, 
                             exp_transform = FALSE,
                             group_race = TRUE) {
  result <- df %>%
    st_drop_geometry() %>% 
    group_by(model, race) %>%
    summarize(
      min = ifelse(exp_transform,
                   min(exp({{ column_name }}), na.rm = TRUE),
                   min({{ column_name }}, na.rm = TRUE)),
      p250 = ifelse(exp_transform,
                    stats::quantile(exp({{ column_name }}), .250, na.rm = TRUE),
                    stats::quantile({{ column_name }}, .250, na.rm = TRUE)),
      p500 = ifelse(exp_transform,
                    stats::median(exp({{ column_name }}), na.rm = TRUE),
                    stats::median({{ column_name }}, na.rm = TRUE)),
      p750 = ifelse(exp_transform,
                    stats::quantile(exp({{ column_name }}), .750, na.rm = TRUE),
                    stats::quantile({{ column_name }}, .750, na.rm = TRUE)),
      max = ifelse(exp_transform,
                   max(exp({{ column_name }}), na.rm = TRUE),
                   max({{ column_name }}, na.rm = TRUE))
    ) %>%
    mutate(param = param_name,
           min = round(min, 2),
           p250 = round(p250, 2),
           p500 = round(p500, 2),
           p750 = round(p750, 2),
           max = round(max, 2))
  return(result)
}

# apply function to get summary stats table
# note: I don't think the first two can come off of FASSE because they show small counts
summary_stats <- rbind(
  # summarize_column(pos_med_sf, 
  #                  column_name = person_years,
  #                  param_name = "person years"),
  # summarize_column(pos_med_sf, 
  #                  column_name = observed,
  #                  param_name = "observed"),
  # summarize_column(df = pos_med_sf, 
  #                  column_name = y, 
  #                  param_name = "Observed/expected", 
  #                  exp_transform = TRUE),
  summarize_column(df = pos_med_sf, 
                   column_name = pos_med_y, 
                   param_name = "Standardized risk (lambda)", 
                   exp_transform = TRUE),
  summarize_column(df = pos_med_sf, 
                   column_name = risk_ratio, 
                   param_name = "Risk ratio",
                   group_race = FALSE)
  )

# make risk ratio a single row (same for both races)
summary_stats %<>%
  filter(!(param == "Risk ratio" & race == 2)) %>%
  mutate(race = ifelse(param == "Risk ratio", "Both", race))

# change race values to white and black
summary_stats %<>%
  mutate(race = ifelse(race == 1, "White", race)) %>%
  mutate(race = ifelse(race == 2, "Black", race)) %>%
  select(c(param, everything()))

# display without scientific notation
summary_stats[] <- lapply(summary_stats, function(x) if(is.numeric(x)) format(x, scientific = FALSE) else x)


# rename columns
summary_stats <- summary_stats |>
  rename(Parameter = param,
         Model = model,
            Race = race,
            Minimum = min,
            `25th percentile` = p250,
            `50th percentile` = p500,
            `75th percentile` = p750,
            Maximum = max)
summary_stats |> pander()

# save table
write_rds(summary_stats, file = paste0("results/tables/summary_stats.rds"))
```

### In a figure

```{r}
# plot distribution of lambdas
pos_med %>% 
  mutate(model = ifelse(model == "m2", "Unadjusted", model),
         model = ifelse(model == "m2", "Adjusted for PIR", model),
         model = ifelse(model == "m3", "Adjusted for PM2.5", model)) %>%
  mutate(model = factor(model, levels = c("Unadjusted", "Adjusted for PIR", "Adjusted for PM2.5"))) %>%
  ggplot() +
  geom_density(aes(x = pos_med_y, col = model)) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(x = expression('Smoothed standardized risk ('*lambda*')'),
       y = "Density",
       col = "") +
  facet_wrap(~race, labeller = as_labeller(race_labels), nrow = 2) +
  theme_light() +
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(color = "black"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank())
```


```{r}
# plot distributions of risk ratios
pos_med %>% 
  mutate(model = ifelse(model == "m2", "Unadjusted", model),
         model = ifelse(model == "m2", "Adjusted for PIR", model),
         model = ifelse(model == "m3", "Adjusted for PM2.5", model)) %>%
  mutate(model = factor(model, levels = c("Unadjusted", "Adjusted for PIR", "Adjusted for PM2.5"))) %>%
  ggplot() +
  geom_density(aes(x = risk_ratio, col = model)) +
  geom_vline(xintercept = 1, lty = 2) +
  labs(x = "Risk Ratio (Black/White)",
       y = "Density",
       col = "") +
  theme_light() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank())
```


## Empirical variances

```{r}
# load list with all results and bind into df
ev <- read_rds(paste0("results/tables/empirical_variance.rds"))
ev <- bind_rows(ev, .id = "model")

# spread columns by model
ev |>
  spread(key = model, value = Estimate) |>
  arrange(Parameter) |>
  rename(`Model 1` = m2,
         `Model 2` = m2,
         `Model 3` = m3) |>
  pander()
```

