
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
```

```{r, message=FALSE, warning=FALSE}
## Load libraries ----
library(tidyverse)
library(magrittr)
library(sf)
library(cowplot)
library(viridis)
library(RColorBrewer)
```

```{r}
## Read data ----
stanfit_samples <- read_rds("data/intermediate/m0-stanfit_samples.rds")
stanfit_summary <- read_rds("data/intermediate/m0-stanfit_summary.rds")
county_sf <- read_rds("data/symlinks/temp/county_sf.rds")
pos_med <- read_rds("data/intermediate/m0-pos_med.rds")

pos_med_sf <- county_sf %>% 
  select(county) %>% 
  right_join(pos_med)

dim(county_sf)
dim(pos_med_sf)
colnames(pos_med_sf)
```

```{r}
county_sf <- read_rds("data/symlinks/temp/county_sf.rds")
county_fips_df <- read_rds("data/symlinks/temp/county_fips_df.rds")
state_sf <- read_sf("data/county/tl_2015_us_state/tl_2015_us_state.shp")
state_sf %<>% 
  left_join(
    county_sf %>% 
      left_join(county_fips_df) %>% 
      st_drop_geometry() %>% 
      distinct(STATEFP, s_idx, st_abbr)
  ) %>% 
  filter(!is.na(s_idx))
```


## Observed vs Predicted

### log ratios (y)
# I think these are the SMRs

```{r}
pos_med_sf %>% 
  mutate(
    res = y - pos_med_y,
    res_is_out = if_else(abs(res) > 2/3, 1, 0) 
  ) %>% 
  ggplot() +
  geom_sf(aes(fill = exp(y), col = as.factor(res_is_out)), lwd = 0.2) +
  facet_grid(~race) +
  theme(legend.position = "bottom") + 
  theme_map()
```


# same thing but without color borders --> those are just showing where SMR and lambda are very different

```{r}
# exponentiate because we already took the log to get y
pos_med_sf %>% 
  ggplot() +
  geom_sf(aes(fill = exp(y)), lwd = 0.00) +
  scale_fill_viridis(option = "D") + 
  facet_grid(~race) +
  theme(legend.position = "bottom") + 
  theme_map()
```

### MAP TEST
# same thing but try using fipschar instead of sf

```{r}
library(urbnmapr)

map_data <- map_data("us")

pos_med_sf %>% 
  ggplot() +
  geom_sf(aes(fill = exp(y)), lwd = 0.00) +
  scale_fill_viridis(option = "D") + 
  facet_grid(~race) +
  theme(legend.position = "bottom") + 
  theme_map()


```

### posterior median log ratios (pos_med_y)

In this model the "prediction" is a smoothed version of the observed.

```{r}
pos_med_sf %>% 
  ggplot() +
  geom_sf(aes(fill = pos_med_y), lwd = 0.00) +
  facet_grid(~race) +
  scale_fill_viridis(option = "D") + 
  theme_map() + 
  theme(legend.position = "bottom") 
```

## Risk decomposition

## state effects

```{r}
param = rownames(stanfit_summary$summary)
nu = tibble(
  param = param[grep("nu", param)],
  med = stanfit_summary$summary[grep("nu", param),"50%"],
  s_idx = 1:length(grep("nu", param))
)

state_sf %>% 
  left_join(nu) %>% 
  ggplot() + 
  geom_sf(aes(fill = med))
```

## shared unscaled

```{r}
summary(pos_med$shared_unscaled)
hist(pos_med$shared_unscaled)
```

```{r}
summary(pos_med$specific)
hist(pos_med$specific)
```

```{r}
# find outliers for phi
pos_med_sf %<>% 
  mutate(
    shared_unscaled_inliers = if_else(shared_unscaled > -1 & shared_unscaled < 1, 
                                      shared_unscaled, as.numeric(NA)),
    shared_unscaled_outliers = if_else(shared_unscaled <= -1 | shared_unscaled >= 1, 
                                       shared_unscaled, as.numeric(NA)),
    shared_unscaled_outliers = if_else(shared_unscaled_outliers <= -1, -1, shared_unscaled_outliers),
    shared_unscaled_outliers = if_else(shared_unscaled_outliers >= 1, 1, shared_unscaled_outliers)
  ) 

pos_med_sf %>% 
  st_drop_geometry() %>% 
  select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# look at outliers
filter(pos_med_sf, !is.na(shared_unscaled_outliers)) %>% 
  st_drop_geometry() %>% 
  select(shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

# Plot three risk surfaces

```{r}
myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = myPalette(100), 
                           limits = c(-1, 1),
                           na.value = "gray70")

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  theme_map() +
  labs(subtitle = "white specific") + 
  theme(legend.title = element_blank()) + 
  sc

# shared unscaled (phi)
p2 <- pos_med_sf %>% 
  filter(race == 1) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared_unscaled_inliers), lwd = 0.0) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  theme_map() + 
  labs(subtitle = "shared unscaled") + 
  sc

# psi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  theme_map() + 
  labs(subtitle = "black specific") +
  sc
  
p <- cowplot::plot_grid(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3  + theme(legend.position = "none"),
  nrow = 3)

cowplot::plot_grid(
  p, 
  cowplot::get_legend(p1), 
  ncol = 2
)
```

## same thing but only significant counties

```{r}
# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  theme_map() +
  labs(subtitle = "white specific") + 
  theme(legend.title = element_blank()) + 
  sc

# shared unscaled (phi)
p2 <- pos_med_sf %>% 
  mutate(shared_unscaled_inliers = ifelse(shared_unscaled_sig, shared_unscaled_inliers, NA)) %>%
  filter(race == 1) %>% #shared_unscaled is non-race specific, so choose one
  ggplot() +
  geom_sf(aes(fill = shared_unscaled_inliers), lwd = 0.0) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  theme_map() + 
  labs(subtitle = "shared unscaled") + 
  sc

# psi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  mutate(specific = ifelse(specific_sig, specific, NA)) %>%
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  theme_map() + 
  labs(subtitle = "black specific") +
  sc
  
p <- cowplot::plot_grid(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3  + theme(legend.position = "none"),
  nrow = 3)

cowplot::plot_grid(
  p, 
  cowplot::get_legend(p1), 
  ncol = 2
)
```

# this looks different from fig 2 in paper because that fig has exponentiated

```{r}
# look at counties with zero person-years
filter(pos_med_sf, person_years == 0) %>% 
  st_drop_geometry() %>% 
  select(county, race, person_years, shared_unscaled, shared_unscaled_inliers, shared_unscaled_outliers)
```

```{r}
# if zero person years, make "specific" and "shared" columns NA
pos_med_sf %<>% 
  mutate(
    specific = if_else(person_years == 0, as.numeric(NA), specific),
    shared_unscaled_inliers = if_else(person_years == 0, as.numeric(NA), shared_unscaled_inliers),
    shared_unscaled_outliers = if_else(person_years == 0, as.numeric(NA), shared_unscaled_outliers)
  )
```

```{r}
# now recreate the same plot but without counties with zero person-years
myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = myPalette(100), 
                           limits = c(-1, 1))

p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  theme_map() +
  labs(subtitle = "white specific") + 
  theme(legend.title = element_blank()) + 
  sc

p2 <- pos_med_sf %>% 
  filter(race == 1) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared_unscaled_inliers), lwd = 0.0) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == 1), color = "red", lwd = 0.5) + 
  geom_sf(data = filter(pos_med_sf, shared_unscaled_outliers == -1), color = "blue", lwd = 0.5) + 
  theme_map() + 
  labs(subtitle = "shared unscaled") + 
  sc

p3 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.0) + 
  theme_map() + 
  labs(subtitle = "black specific") +
  sc
  
p <- cowplot::plot_grid(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3  + theme(legend.position = "none"),
  nrow = 3)

cowplot::plot_grid(
  p, 
  cowplot::get_legend(p1), 
  ncol = 2
)
```

## shared unscaled

```{r}
# now split the middle map into scaled phi for each race

shared_ <- c(pos_med$shared, pos_med$specific)
myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = myPalette(100), 
                           limits = c(min(shared_), max(shared_)))

# psi for white people
p1 <- pos_med_sf %>% 
  filter(race == 1) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.00) + 
  theme_map() +
  labs(subtitle = "white specific") + 
  theme(legend.title = element_blank()) + 
  sc

# scaled phi for white people
p2 <- pos_med_sf %>% 
  filter(race == 1) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared), lwd = 0.00) + 
  theme_map() + 
  labs(subtitle = "shared white") + 
  sc

# scaled phi for black people
p3 <- pos_med_sf %>% 
  filter(race == 2) %>% #shared_unscaled is non-race specific
  ggplot() +
  geom_sf(aes(fill = shared), lwd = 0.00) + 
  theme_map() + 
  labs(subtitle = "shared black") + 
  sc

# psi for black people
p4 <- pos_med_sf %>% 
  filter(race == 2) %>% 
  ggplot() +
  geom_sf(aes(fill = specific), lwd = 0.00) + 
  theme_map() + 
  labs(subtitle = "black specific") +
  sc
  
p <- cowplot::plot_grid(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3  + theme(legend.position = "none"),
  p4  + theme(legend.position = "none"),
  nrow = 2, 
  ncol = 2)

cowplot::plot_grid(
  p, 
  cowplot::get_legend(p1), 
  ncol = 2, 
  rel_widths = c(0.8, 0.2)
)
```


#############################################################
#############################################################
#############################################################

## New maps from Lauren!

```{r}
# recreate figure 2

```

```{r}
# recreate figure 3

# plot distribution of risk ratios
pos_med_sf %>%
  ggplot() +
  geom_density(aes(risk_ratio))
```

```{r}
# map of risk ratios

# need to change color scale to show <1 and >1
myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = myPalette(100), 
                           limits = c(min(pos_med_sf$risk_ratio), max(pos_med_sf$risk_ratio)))

pos_med_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA))
  ) %>%
  select(county, geometry, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio), lwd = 0.00) + 
  labs(fill = "Risk ratio") +
  theme_map() +
  sc
```


