
```{r setup, include=FALSE}
## Set working directory as project root
library(here)
knitr::opts_knit$set(root.dir = here())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, message=FALSE, warning=FALSE}
## Load libraries ----
library(tidyverse)
library(magrittr)
library(sf)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(pander)
library(data.table)
library(ggthemes)
library(knitr)
library(kableExtra)
library(xtable)
```

```{r}
# set path for saving figures
fig_path <- "results/figures/final/"
```

## Load data

```{r}
# load list with all model results and bind into df
hosp_res <- read_rds("data/intermediate/hosp-dual_model_results.rds")
hosp_res <- bind_rows(hosp_res, .id = "model")

# # limiting to model 1 for now
# hosp_res <- hosp_res %>%
#   filter(model == "hosp_m1")
```

```{r}
## Read data ----
hosp_ratios_df_comb <- read_rds('data/symlinks/scratch/hosp_combined_data_dual.rds')
hosp_ratios_df <- read_rds('data/symlinks/scratch/hosp_ratios_df_dual.rds')
county_sf <- read_rds("data/symlinks/scratch/county_sf.rds")
county_sf_zero_b <- read_rds("data/symlinks/scratch/county_sf_zero_b.rds")

hosp_res_sf <- county_sf %>% 
  select(county) %>% 
  right_join(hosp_ratios_df_comb) %>%
  right_join(hosp_ratios_df) %>%
  right_join(hosp_res)

colnames(hosp_res_sf)
```

```{r}
county_fips_df <- read_rds("data/symlinks/scratch/county_fips_df.rds")
state_sf <- read_sf("data/county/cb_2015_us_state/cb_2015_us_state_20m.shp")
state_sf %<>% 
  left_join(
    county_sf %>% 
      left_join(county_fips_df) %>% 
      st_drop_geometry() %>% 
      distinct(STATEFP, s_idx)
  ) %>% 
  filter(!is.na(s_idx))
```

```{r}
# change projection to Albers Equal Area Conic (good for maps of US)

hosp_res_sf <- st_transform(hosp_res_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
state_sf <- st_transform(state_sf, 
                           crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
county_sf <- st_transform(county_sf,
                                 crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
county_sf_zero_b <- st_transform(county_sf_zero_b,
                                 crs = st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5
                                        +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"))
```

```{r}
# race labels for plotting
race_labels <- c(
  `1` = "White",
  `2` = "Black"
)

# model labels for plotting
model_labels <- c(
  hosp_m1 = "Unadjusted",
  hosp_m2 = "Adjusted for\n price-to-income ratio",
  hosp_m3 = "Adjusted for PM\u2082.\u2085"
)

# race and model labels for plotting
race_model_labels <- c(
  `1` = "White",
  `2` = "Black",
  hosp_m1 = "Unadjusted",
  hosp_m2 = "Adjusted for\n price-to-income ratio",
  hosp_m3 = "Adjusted for PM\u2082.\u2085"
)
```

```{r}
# make a new column for log(observed/expected) separated by race
hosp_res_sf %<>%
  mutate(y_stage1 = log(observed/expected_stage1),
         y_stage2 = log(observed/expected_stage2))

# make a new column for log(observed/expected) combined race
hosp_res_sf %<>%
  mutate(y_stage1_total = log(observed_total/expected_stage1_total))

# make a new column for rate of hospitalization (observed/person-years)
hosp_res_sf %<>% 
  mutate(hosp_rate_obs = observed/person_years)

# and a new column for expected rate of hospitalization (expected/person-years)
hosp_res_sf %<>% 
  mutate(hosp_rate_exp_stage1 = expected_stage1/person_years,
         hosp_rate_exp_stage2 = expected_stage2/person_years)
```


## Understand the data/covariates

```{r}
# #png(file = paste0(fig_path, "pct_black.png"), width = 800, height = 500)
# 
# # plot % of person-years contributed by black people in each county
# hosp_res_sf %>%
#   filter(model == "hosp_m1") %>%
#   group_by(county) %>%
#   summarise(pct_blk = 100 * (person_years[race == 2] /
#                                (person_years[race == 1] + person_years[race == 2]))) %>%
#   ggplot() +
#   geom_sf(aes(fill = pct_blk), lwd = 0.0) +
#   geom_sf(data = county_sf_zero_b, fill = "#440154", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.2) +
#   labs(fill = "% of person-years \ncontributed by \nBlack adults") +
#   scale_fill_viridis(option = "D",
#                      limits = c(0, 70)) +
#   theme_map() +
#   theme(legend.title = element_text(size = 14),
#         legend.text = element_text(size = 14),
#         legend.position = "right",
#         plot.background = element_rect(fill = "white", color = NA))
# 
# ggsave(paste0(fig_path, "pct_black.png"), width = 8, height = 5, dpi = 300)
# 
# #dev.off()
```

```{r}
# # map PM2.5, HPIR, median household income
#   
# #png(file = paste0(fig_path, "covariate_maps.png"), width = 800, height = 500)
# 
# pir_plot <- hosp_res_sf %>%
#   filter(model == "hosp_m1") %>%
#   ggplot() +
#   geom_sf(aes(fill = log10(house_price_income_ratio)), alpha = 1, lwd = 0.0) +
#   labs(fill = "log (HPIR)") +
#   scale_fill_viridis(option = "D",
#                      guide = guide_colorbar(barwidth = 10, barheight = 1)) +
#   geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   theme_map() +
#   theme(legend.position = "bottom",
#         plot.background = element_rect(fill = "white", color = NA),
#         panel.background = element_blank(),
#         plot.margin = margin(0, 0, 0, 0))
# 
# # hosp_res_sf %>%
# #   filter(model == "hosp_m1") %>%
# #   ggplot() +
# #   geom_sf(aes(fill = med_house_income), alpha = 1, lwd = 0.0) +
# #   labs(fill = "Median household \nincome") +
# #   scale_fill_viridis(option = "D") +
# #   geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
# #   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
# #   theme_map() +
# #   theme(legend.position = "right")
# 
# pm25_plot <- hosp_res_sf %>%
#   filter(model == "hosp_m1") %>%
#   ggplot() +
#   geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
#   geom_sf(aes(fill = pm25), alpha = 1, lwd = 0.0) +
#   labs(fill = expression('PM'[2.5]*' (\u00B5g/m'^3*')')) +
#   scale_fill_viridis(option = "D",
#                      guide = guide_colorbar(barwidth = 10, barheight = 1)) +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   theme_map() +
#   theme(legend.position = "bottom",
#         plot.background = element_rect(fill = "white", color = NA),
#         panel.background = element_blank(),
#         plot.margin = margin(0, 0, 0, 0))
# 
# cov_maps <- ggarrange(pir_plot, pm25_plot) +
#   theme(plot.background = element_rect(fill = "white", color = NA))
# 
# cov_maps
# 
# #ggarrange(pir_plot, pm25_plot)
# 
# # library(gridExtra)
# # grid.arrange(pir_plot, pm25_plot)
# 
# ggsave(paste0(fig_path, "covariate_maps.png"), width = 8, height = 3, dpi = 300)
# 
# #dev.off()
```

```{r}
# # number of person-years observed in each county by race
# hosp_res_sf %>%
#   filter(model == "hosp_m1") %>%
#   ggplot() +
#   geom_sf(aes(fill = log(person_years)), alpha = 1, lwd = 0.0) +
#   labs(fill = "log(person-years)") +
#   scale_fill_viridis(option = "D") +
#   geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   facet_grid(~race, 
#              labeller = as_labeller(race_labels)) +
#   theme_map() +
#   theme(legend.position = "bottom",
#         strip.background = element_blank())
```

```{r}
# # which counties have very few person-years?
# hosp_res_sf %>%
#   filter(model == "hosp_m1" & race == "2") %>%
#   mutate(few_person_years = ifelse(person_years < 100, 1, 0)) %>%
#   mutate(few_person_years = ifelse(person_years > 0, few_person_years, NA)) %>%
#   ggplot() +
#   geom_sf(aes(fill = as.factor(few_person_years)), alpha = 1, lwd = 0.03) +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   labs(fill = "< 100 Black person-years") +
#   # facet_grid(~race, 
#   #            labeller = as_labeller(race_labels)) +
#   theme(legend.position = "bottom") +
#   theme_map()
# 
# hosp_res_sf %>%
#   filter(model == "hosp_m1" & race == "2") %>%
#   mutate(few_person_years = ifelse(person_years < 100, 1, 0)) %>%
#   mutate(few_person_years = ifelse(person_years > 0, few_person_years, NA)) %>%
#      count(few_person_years)
```


## Pre-modeling visualizations of observed and expected

### Expected rates table

```{r}
# exp_rates_1 <- read_rds(paste0("results/tables/hosp_expected_rates_race-specific_dual.rds"))
# exp_rates_2 <- read_rds(paste0("results/tables/hosp_expected_rates_overall_dual.rds"))
# 
# exp_rates_tab <- left_join(exp_rates_1, exp_rates_2)
# 
# # get table in latex
# print(xtable(exp_rates_tab, type = "latex"), file = "results/tables/final/hosp_national_rates_dual.tex")

# exp_rates_1 %>% pander()
# exp_rates_2 %>% pander()

# # Create a basic table with kable
# basic_table <- kable(exp_rates_tab, 
#                      format = "html", 
#                      align = "c", 
#                      caption = "Age- and sex- standardized national rates of hospitalization with diagnosis codes related to ADRD (# of hospitalizations per person-year).")
# 
# # Customize the table appearance with kableExtra
# styled_table <- basic_table %>%
#   kable_styling(full_width = FALSE, bootstrap_options = c("striped"))
# styled_table
```


### Observed rate of hospitalization with ADRD (observed/person-years)

```{r}
# try capping at 12%--otherwise we're just seeing really small counties
# alternatively, could restrict this to counties with a certain number of person-years observed

#png(file = paste0(fig_path, "hosp_rate_obs_dual.png"), width = 900, height = 400)

hosp_res_sf %>%
  filter(model == "hosp_m1") %>%
  mutate(hosp_rate_obs = ifelse(hosp_rate_obs > 0.12, 0.12, hosp_rate_obs)) |>
  mutate(hosp_rate_obs = ifelse(observed <= 10, NA, hosp_rate_obs)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_obs * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Rate of hosp. \nwith ADRD (%)") +
  scale_fill_viridis(option = "D", na.value = "gray70") +
  geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.text = element_text(size = 14))

#dev.off()

# could also try lumping together white and black (this just tells us where there are more ADRD hospitalizations)
```


### Expected rate of hospitalization with ADRD (expected/person-years)

```{r}
# try capping at 5%--otherwise we're just seeing really small counties
# alternatively, could restrict this to counties with a certain number of person-years
hosp_res_sf %>%
  filter(model == "hosp_m1") %>%
  mutate(hosp_rate_exp_stage1 = ifelse(hosp_rate_exp_stage1 > 0.12, 
                                       0.12, hosp_rate_exp_stage1)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_exp_stage1 * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Expected rate of \nhosp. with ADRD (%), \nstage 1") +
  scale_fill_viridis(option = "D") +
  geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() +
  theme(legend.position = "bottom",
        strip.background = element_blank())

hosp_res_sf %>%
  filter(model == "hosp_m1") %>%
  mutate(hosp_rate_exp_stage2 = ifelse(hosp_rate_exp_stage2 > 0.12, 
                                       0.12, hosp_rate_exp_stage2)) |>
  ggplot() +
  geom_sf(aes(fill = hosp_rate_exp_stage2 * 100), alpha = 1, lwd = 0.0) +
  labs(fill = "Expected rate of \nhosp. with ADRD (%), \nstage 2") +
  scale_fill_viridis(option = "D") +
  geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() +
  theme(legend.position = "bottom",
        strip.background = element_blank())

# could also try lumping together white and black (this just tells us where there are more ADRD hospitalizations)
```

### Observed vs. expected

```{r}
# map of observed counts
hosp_res_sf %>%
  ggplot() +
  geom_sf(aes(fill = observed), lwd = 0.03) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = "Observed counts") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() + 
  theme(legend.position = "bottom",
        strip.background = element_blank())

# map of expected counts (stage 1)
hosp_res_sf %>%
  ggplot() +
  geom_sf(aes(fill = expected_stage1), lwd = 0.03) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  labs(fill = "Expected counts (stage 1)") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() + 
  theme(legend.position = "bottom",
        strip.background = element_blank())

# map of expected counts (stage 2)
hosp_res_sf %>%
  ggplot() +
  geom_sf(aes(fill = expected_stage2), lwd = 0.03) +
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  scale_fill_viridis(option = "D", labels = scales::comma) +
  labs(fill = "Expected counts (stage 2)") +
  facet_grid(~race, labeller = as_labeller(race_labels)) +
  theme_map() + 
  theme(legend.position = "bottom",
        strip.background = element_blank())
```


## Model results

### Stage 1 raw vs. smoothed lambdas

```{r}

# Note: y is log(observed) - log(expected) or log(observed/expected), so exp(y) is observed/expected. exp(y) is analogous to the standardized mortality ratio (SMR) in Kiang et al.

# note: it's better to plot y than log(y)!
# if zero hosp are observed and 1 is expected, log(y) = -Inf, but y = 0
# (if zero hosp are expected, then y and log(y) will both be undefined)

# gather raw observed/expected and lambdas
stage1_total_risk_df <- hosp_res_sf |>
  filter(race == 1) |>
  select(county, geometry, model, 
         y_stage1_total, log_lambda_stage1) |>
  # gather raw and modeled lambdas
  gather(key = "raw_vs_model", value = "estimate", y_stage1_total, log_lambda_stage1) %>%
  # separate category for each model
  mutate(raw_vs_model = ifelse(raw_vs_model == "log_lambda_stage1", model, "raw")) %>%
  select(-model) %>%
  distinct()

# set levels for plotting
stage1_total_risk_df$raw_vs_model <- factor(stage1_total_risk_df$raw_vs_model,
                                       levels = c("raw", "hosp_m1", "hosp_m2", "hosp_m3"))

# labels for facets
stage1_total_risk_labels <- c(
  raw = "Raw observed/expected",
  #hosp_m1 = "Smoothed (\u03BB)", # CHANGE IF LOOKING AT COVARIATES
  hosp_m1 = "Unadjusted", # CHANGE IF LOOKING AT COVARIATES
  hosp_m2 = "Adjusted for\n price-to-income ratio",
  hosp_m3 = "Adjusted for PM\u2082.\u2085",
  "1_specific" = "White-specific\n (\u03C8\u2081)",
  "both_shared_unscaled" = "Unscaled shared\n (\u03A6)",
  "2_specific" = "Black-specific\n (\u03C8\u2082)"
)

# plot
stage1_total_risk_df |>
  ggplot() +
  geom_sf(aes(fill = exp(estimate)), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray60", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = expression('Standardized risk')) +
  scale_fill_viridis(option = "A", 
                     na.value = "gray60",
                     limits = c(0, 3)) +
  facet_grid(~raw_vs_model,
             labeller = as_labeller(stage1_total_risk_labels)
             ) +
  theme_map() + 
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.position = "bottom",
        strip.text = element_text(size = 14),
        strip.background = element_blank())

```

```{r}
# what if we use observed vs. expected for stage 1, but plot separately by race?

# plot
hosp_res_sf |>
  ggplot() +
  geom_sf(aes(fill = exp(y_stage1)), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  labs(fill = expression('Standardized risk')) +
  scale_fill_viridis(option = "A", 
                     na.value = "gray60",
                     limits = c(0, 3)) +
  facet_grid(model~race,
             labeller = as_labeller(race_model_labels)
             ) +
  theme_map() + 
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        strip.text = element_text(size = 10),
        strip.background = element_blank())

```




### Stage 2 raw vs. smoothed lambdas

```{r}
# set up data

# gather raw observed/expected and lambdas
stage2_total_risk_df <- hosp_res_sf |>
  select(county, geometry, model, race,
         y_stage2, log_lambda_stage2) |>
  # gather raw and modeled lambdas
  gather(key = "raw_vs_model", value = "estimate", y_stage2, log_lambda_stage2) %>%
  # separate category for each model
  mutate(raw_vs_model = ifelse(raw_vs_model == "log_lambda_stage2", model, "raw")) %>%
  select(-model) %>%
  distinct()

# set levels for plotting
stage2_total_risk_df$raw_vs_model <- factor(stage2_total_risk_df$raw_vs_model,
                                       levels = c("raw", "hosp_m1", "hosp_m2", "hosp_m3"))
```


```{r}
#--- figure with only the unadjusted model

# labels for facets
stage2_total_risk_1m_labels <- c(
  `1` = "White",
  `2` = "Black",
  raw = "Raw\n observed/expected",
  hosp_m1 = "Smoothed\n observed/expected (\u03BB)",
  "1_specific" = "White-specific\n (\u03C8\u2081)",
  "both_shared_unscaled" = "Unscaled shared\n (\u03A6)",
  "2_specific" = "Black-specific\n (\u03C8\u2082)"
)

#png(file = paste0(fig_path, "hosp_raw_vs_lambda_stage2_1m.png"), width = 700, height = 400)

# plot
stage2_total_risk_df |>
  filter(raw_vs_model == "raw" | raw_vs_model == "hosp_m1") |>
  mutate(estimate = ifelse(estimate >= log(3), log(3)-0.001, estimate)) |>
  ggplot() +
  geom_sf(aes(fill = exp(estimate)), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  theme_map() + 
  labs(fill = expression('Standardized risk')) +
  facet_grid(race ~ raw_vs_model,
             labeller = as_labeller(stage2_total_risk_1m_labels)
             ) +
  theme(legend.title = element_text(size = 12, vjust = 0.8),
        legend.text = element_text(size = 9),
        legend.position = "bottom",
        strip.text = element_text(size = 12),
        strip.background = element_blank(),
        plot.background = element_rect(fill = "white", color = NA)) +
  # this scale is capping at max in previous plot
  scale_fill_viridis(option = "A", 
                     na.value = "gray70",
                     limits = c(0, 3),
                     labels = c("0", "1", "2", "3+"),
                     guide = guide_colorbar(ticks = FALSE,
                                            barwidth = 12, barheight = 0.5))

ggsave(paste0(fig_path, "hosp_raw_vs_lambda_stage2_1m_dual.png"), width = 6, height = 4, dpi = 300)

#dev.off()
```

```{r}
# same thing but remove raw maps (not needed for clinical audience)

# labels for facets
stage2_no_raw_labels <- c(
  `1` = "White",
  `2` = "Black",
  raw = "Raw observed/expected",
  hosp_m1 = "Smoothed (\u03BB)",
  "1_specific" = "White-specific\n (\u03C8\u2081)",
  "both_shared_unscaled" = "Unscaled shared\n (\u03A6)",
  "2_specific" = "Black-specific\n (\u03C8\u2082)"
)

#png(file = paste0(fig_path, "hosp_lambda_stage2_1m.png"), width = 700, height = 300)

# plot
stage2_total_risk_df |>
  filter(raw_vs_model == "raw" | raw_vs_model == "hosp_m1" | raw_vs_model == "raw") |>
  mutate(estimate = ifelse(estimate >= log(3), log(3)-0.001, estimate)) |>
  ggplot() +
  geom_sf(aes(fill = exp(estimate)), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
  theme_map() + 
  labs(fill = expression('Standardized risk')) +
  facet_grid(~race,
             labeller = as_labeller(stage2_no_raw_labels)
             ) +
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.position = "bottom",
        strip.text = element_text(size = 16),
        strip.background = element_blank()) +
  # this scale is capping at max in previous plot
  scale_fill_viridis(option = "A", 
                     na.value = "gray70",
                     limits = c(0, 3),
                     labels = c("0", "1", "2", "3+")) + 
  guides(fill = guide_colorbar(title.position = "top", barwidth = 10))

dev.off()

```



```{r}
#--- figure with all three models

# # labels for facets
# stage2_total_risk_3m_labels <- c(
#   `1` = "White",
#   `2` = "Black",
#   raw = "Raw observed/expected",
#   hosp_m1 = "Unadjusted", # CHANGE IF USING COVARIATES
#   hosp_m2 = "Adjusted for\n price-to-income ratio",
#   hosp_m3 = "Adjusted for PM\u2082.\u2085",
#   "1_specific" = "White-specific\n (\u03C8\u2081)",
#   "both_shared_unscaled" = "Unscaled shared\n (\u03A6)",
#   "2_specific" = "Black-specific\n (\u03C8\u2082)"
# )
# 
# png(file = paste0(fig_path, "hosp_raw_vs_lambda_stage2_3m.png"), width = 700, height = 400)
# 
# # plot
# stage2_total_risk_df |>
#   mutate(estimate = ifelse(estimate >= log(3), log(3)-0.001, estimate)) |>
#   ggplot() +
#   geom_sf(aes(fill = exp(estimate)), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   theme_map() + 
#   labs(fill = expression('Standardized risk')) +
#   facet_grid(race ~ raw_vs_model,
#              labeller = as_labeller(stage2_total_risk_3m_labels)
#              ) +
#   theme(legend.title = element_text(size = 14),
#         legend.text = element_text(size = 12),
#         legend.position = "bottom",
#         strip.text = element_text(size = 12),
#         strip.background = element_blank()) +
#   # this scale is capping at max in previous plot
#   scale_fill_viridis(option = "A", 
#                      na.value = "gray70",
#                      limits = c(0, 3),
#                      labels = c("0", "1", "2", "3+"))
# 
# dev.off()
```

```{r}
# # plot only smoothed results (for poster)
# 
# long_race_labels <- c(
#   `1` = "Non-Hispanic White",
#   `2` = "Non-Hispanic Black"
# )
# 
# # plot
# hosp_res_sf %>%
#   mutate(log_lambda_stage2 = ifelse(log_lambda_stage2 > log(3), log(3), log_lambda_stage2)) %>%
#   ggplot() +
#   geom_sf(aes(fill = exp(log_lambda_stage2)), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray70", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.2) +
#   theme_map() + 
#   labs(fill = expression('Smoothed standardized risk ('* lambda * ')')) +
#   facet_grid(~ race,
#              labeller = as_labeller(long_race_labels)
#              ) +
#   theme(legend.title = element_text(size = 14),
#         legend.text = element_text(size = 14),
#         legend.position = "bottom",
#         strip.text = element_text(size = 16),
#         strip.background = element_blank()) +
#   # this scale is capping at max in previous plot
#   scale_fill_viridis(option = "A", 
#                      na.value = "gray60")
```



```{r}
# now compare raw ratios (observed/expected--unsmoothed) to model results (lambda--smoothed)

# stage 1
sc <- scale_fill_gradient2(high = "red", mid = "white", low = "blue", na.value = "gray50",
                         midpoint = 0)
hosp_res_sf %>%
  mutate(
    res = y_stage1 - log_lambda_stage1#,
    #res_is_out = if_else(abs(res) > 1, "different", "similar")
  ) %>%
  ggplot() +
  geom_sf(aes(fill = res 
              #fill = as.factor(res_is_out)), lwd = 0.1
    )) +
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Raw ratio - lambda") +
  facet_grid(~ model,
             labeller = as_labeller(model_labels)
             ) +
  theme_map() +
  theme(legend.position = "right",
        strip.background = element_blank()) +
  sc

# stage 2
hosp_res_sf %>%
  mutate(
    res = y_stage2 - log_lambda_stage2,
    res_is_out = if_else(abs(res) > 2/3, "different", "similar")
  ) %>%
  ggplot() +
  geom_sf(aes(#fill = exp(y), 
              fill = as.factor(res_is_out)), lwd = 0.1) +
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Raw ratio & \nlambda are:") +
  facet_grid(race ~ model,
             labeller = as_labeller(race_model_labels)
             ) +
  theme_map() +
  theme(legend.position = "right",
        strip.background = element_blank())
# makes sense that they are more different for Black individuals (more to smooth)
```


### State effects

```{r}
# plot state intercepts (eta)
# note: eta was referred to as nu in the original code!

# param = rownames(stanfit_summary$summary)
# nu = tibble(
#   param = param[grep("nu", param)],
#   med = stanfit_summary$summary[grep("nu", param),"50%"],
#   s_idx = 1:length(grep("nu", param))
# )
# 
# state_sf %>% 
#   left_join(nu) %>% 
#   ggplot() + 
#   geom_sf(aes(fill = med), col = "white") +
#   geom_sf(data = state_sf, fill = NA, col = "white", lwd = 0.1) +
#   labs(fill = expression('State intercept ('*eta*')')) +
#   theme_map()
```

### White-specific, Black-specific, and shared risks (three risk surfaces)

```{r}
# summary(pos_med$shared_unscaled)
# hist(pos_med$shared_unscaled)
```

```{r}
# summary(pos_med$specific)
# hist(pos_med$specific)
# hist(exp(pos_med$specific))
```

#### First, include non-significant counties and use a continuous color scale

```{r}
# set up data

# gather shared and specific columns
shared_specific_df <- hosp_res_sf |>
  select(county, geometry, model, race, 
         phi, phi_sig, psi, psi_sig) |>
  gather(key = "shared_vs_spec", value = "estimate", phi, psi) %>%
  # new sig column for estimate
  mutate(estimate_sig = ifelse(shared_vs_spec == "psi", psi_sig, phi_sig)) |>
  # remove other sig columns for clarity
  select(-c(phi_sig, psi_sig)) |>
  # change race for shared rows
  mutate(race = ifelse(shared_vs_spec == "phi", "both", race)) %>%
  distinct() |>
  mutate(parameter = paste0(race, "_", shared_vs_spec)) 

# set levels for plotting
shared_specific_df$parameter <- factor(shared_specific_df$parameter,
                                       levels = c("1_psi", "both_phi", "2_psi"))

sc <- scale_fill_gradient2(high = "#7600bc", mid = "white", low = "#028A0F", na.value = "gray50",
                         midpoint = 0,
                         guide = guide_colorbar(barwidth = 10, barheight = 1))
```


```{r}
# just one model

# labels for facets
shared_specific_labels_1m <- c(
  "1_psi" = "White-specific\n (\u03C8\u2081)",
  #"1_psi" = "White-specific",
  "both_phi" = "Shared\n (\u03A6)",
  #"both_phi" = "Shared",
  "2_psi" = "Black-specific\n (\u03C8\u2082)"
  #"2_psi" = "Black-specific"
)

# plot
shared_specific_df |>
  filter(model == "hosp_m1") |>
  # only plot significant counties
  # mutate(estimate = ifelse(estimate_sig, estimate, NA)) |>
  ggplot() +
  geom_sf(aes(fill = estimate), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray50", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Risk component") +
  facet_grid(~ parameter,
             labeller = as_labeller(shared_specific_labels_1m)
             ) +
  theme_map() + 
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.position = "bottom",
        strip.text = element_text(size = 14),
        strip.background = element_blank()) +
  # theme(legend.title = element_text(size = 14),
  #       legend.text = element_text(size = 14),
  #       legend.position = "bottom",
  #       strip.text = element_text(size = 16),
  #       strip.background = element_blank()) +
  sc

```

```{r}
# # all three models
# 
# # labels for facets
# shared_specific_labels_3m <- c(
#   hosp_m1 = "Unadjusted",
#   hosp_m2 = "Adjusted for\n price-to-income ratio",
#   hosp_m3 = "Adjusted for PM\u2082.\u2085",
#   #"1_psi" = "White-specific\n (\u03C8\u2081)",
#   "1_psi" = "White-specific",
#   #"both_phi" = "Shared\n (\u03A6)",
#   "both_phi" = "Shared",
#   #"2_psi" = "Black-specific\n (\u03C8\u2082)"
#   "2_psi" = "Black-specific"
# )
# 
# png(file = paste0(fig_path, "hosp_shared_vs_specific_3m.png"), width = 600, height = 400)
# 
# # plot
# shared_specific_df |>
#   # only plot significant counties
#   # mutate(estimate = ifelse(estimate_sig, estimate, NA)) |>
#   ggplot() +
#   geom_sf(aes(fill = estimate), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray50", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   labs(fill = "Risk component") +
#   facet_grid(parameter ~ model,
#              labeller = as_labeller(shared_specific_labels_3m)
#              ) +
#   theme_map() + 
#   theme(legend.title = element_text(size = 12),
#         legend.text = element_text(size = 12),
#         legend.position = "bottom",
#         strip.text = element_text(size = 14),
#         strip.background = element_blank()) +
#   # theme(legend.title = element_text(size = 14),
#   #       legend.text = element_text(size = 14),
#   #       legend.position = "bottom",
#   #       strip.text = element_text(size = 16),
#   #       strip.background = element_blank()) +
#   sc
# 
# dev.off()
```

#### Now, just plot risk as binary and remove non-significant counties

```{r}
# set up data
shared_specific_df %<>%
  mutate(estimate_binary = ifelse(estimate > 0, "High risk", "Low risk")) %>%
  mutate(estimate_binary = ifelse(estimate_sig, estimate_binary, NA))
```


```{r}
# just unadjusted model

# plot
shared_specific_df |>
  filter(model == "hosp_m1") |>
  ggplot() +
  geom_sf(aes(fill = estimate_binary), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray90", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "Risk component:") +
  facet_grid(~ parameter,
             labeller = as_labeller(shared_specific_labels_1m)
             ) +
  theme_map() + 
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "bottom",
        strip.text = element_text(size = 12),
        strip.background = element_blank(),
        plot.background = element_rect(fill = "white", color = NA)) +
  scale_fill_manual(values = c("Low risk" = "#006400", 
                               "High risk" = "#B47EE5"),
                     na.value = "gray90")

ggsave(paste0(fig_path, "hosp_shared_vs_specific_binary_1m_dual.png"), width = 7, height = 2.5, dpi = 300)
```

```{r}
# all three models

# png(file = paste0(fig_path, "hosp_shared_vs_specific_binary_3m.png"), width = 700, height = 500)
# 
# # plot
# shared_specific_df |>
#   ggplot() +
#   geom_sf(aes(fill = estimate_binary), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray90", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   labs(fill = "Risk component:") +
#   facet_grid(parameter ~ model,
#              labeller = as_labeller(shared_specific_labels_3m)
#              ) +
#   theme_map() + 
#   theme(legend.title = element_text(size = 12),
#         legend.text = element_text(size = 12),
#         legend.position = "bottom",
#         strip.text = element_text(size = 10),
#         strip.background = element_blank()) +
#   scale_fill_manual(values = c("Low risk" = "#006400", 
#                                "High risk" = "#B47EE5"),
#                      na.value = "gray90")
# 
# dev.off()
```


### Risk ratios

```{r}
# dichotomize risk ratio--where is white vs. black risk higher?

# make a table--what % of counties are significantly <1? significantly >1?
RR_tab <- hosp_res_sf %>%
  #slice_sample(prop = 0.05) %>%
  filter(race == 2) %>% # same risk ratio for white and black rows
  # select(county, geometry, risk_ratio, risk_ratio_sig) %>%
  # distinct() %>%
  select(-geometry) %>%
  as.data.frame() %>%
  mutate(
    `Risk Ratio` = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    `Risk Ratio` = ifelse(`Risk Ratio` < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  group_by(model) %>%
  count(`Risk Ratio`, name = "Count") %>%
  group_by(model) %>%
  mutate(pcnt = Count / sum(Count) * 100)

# print table
RR_tab |> pander()

```

```{r}
# same table as above BUT black py > 700

# make a table--what % of counties are significantly <1? significantly >1?
RR_tab <- hosp_res_sf %>%
  #slice_sample(prop = 0.05) %>%
  
  filter(race == 2 & person_years > 700) %>% # same risk ratio for white and black rows
  
  # select(county, geometry, risk_ratio, risk_ratio_sig) %>%
  # distinct() %>%
  select(-geometry) %>%
  as.data.frame() %>%
  mutate(
    `Risk Ratio` = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    `Risk Ratio` = ifelse(`Risk Ratio` < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  group_by(model) %>%
  count(`Risk Ratio`, name = "Count") %>%
  group_by(model) %>%
  mutate(pcnt = Count / sum(Count) * 100)

# print table
RR_tab |> pander()
```


```{r}
# plot dichotomized risk ratio
hosp_res_sf %>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, NA),
    risk_ratio = ifelse(risk_ratio < 1, "Risk ratio < 1", "Risk ratio > 1")
  ) %>%
  select(county, geometry, model, risk_ratio) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = factor(risk_ratio)), lwd = 0.06) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
  labs(fill = "") +
  facet_grid(~ model,
             labeller = as_labeller(model_labels)
             ) +
  scale_fill_manual(values = c("Risk ratio < 1" = "#2171b5", 
                               "Risk ratio > 1" = "#fdae6b"),
                    na.value = "gray90") +
  theme_map() +
  theme(strip.background = element_blank())
```


```{r}
# some attempts at continuous color scales (challenging!)

# library(paletteer)
# library(ggthemes)

# # this works, but the blue doesn't get dark enough
# sc <- scale_fill_gradientn(colors = rev(paletteer_c("ggthemes::Red-Blue-White Diverging", 
#                                                     20)[c(1:10, 14)]),
#                            limits = c(min(hosp_res_sf$risk_ratio), max(hosp_res_sf$risk_ratio)),
#                            na.value = "gray70")

# # this is blue to brown for some reason
# library(colorspace)
# sc <- colorspace::scale_fill_continuous_diverging(palette = "Green", 
#                                            na.value = "gray70",
#                                            mid = 1#,
#                                            #limits = c(0.7, 2.5),
#                                            #breaks = c(seq(0.7, 1, by = 0.1), seq(1.1, 2.5, by = 0.2))
#                                            )

# # this works, but again, purple doesn't get dark enough
# sc <- scale_fill_gradient2(low = "#710193", mid = "white", high = "darkgreen",
#                          midpoint = 1)

# Error: Continuous value supplied to discrete scale
# sc <- ggplot2::scale_fill_brewer(NULL, palette = "PRGn", direction = -1)

# # map of risk ratios
# hosp_res_sf %>% 
#   mutate(
#     risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA))
#   ) %>%
#   select(county, geometry, risk_ratio) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio), lwd = 0.1) + 
#   lakes_states_gg +
#   labs(fill = "Risk ratio") +
#   theme_map() +
#   sc
```

```{r}
# try plotting risk ratio with discrete bins (instead of continuous or dichotomous)

# set up data

# set boundaries for each discrete bin
# may need to change!! should be consistent across models and include all RR
# bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2.2, 3)
#bin_breaks <- c(0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, 2.2, 2.4, 3.6)
bin_breaks <- c(0, 0.2, 0.4, 0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, 2.2, 2.4, 3.6)


# get bin labels from bin breaks (get mean)
bin_df <- embed(bin_breaks, 2) |> as.data.frame()
bin_means <- rowMeans(bin_df) |> format(nsmall = 1)

# new column for risk ratio
hosp_res_sf %<>% 
  mutate(
    risk_ratio = ifelse(risk_ratio_sig, risk_ratio, as.numeric(NA)),
    risk_ratio_bins = cut(risk_ratio, 
                          breaks = bin_breaks,
                          labels = bin_means
                          ))

```

```{r}
# save plot for unadjusted model only

hosp_risk_ratios_map_1m <- hosp_res_sf %>% 
  filter(model == "hosp_m1") |>
  select(county, geometry, model, risk_ratio_bins) %>%
  distinct() %>%
  ggplot() +
  geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
  geom_sf(data = county_sf_zero_b, fill = "gray40", lwd = 0.0) + 
  geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.2) +
  labs(fill = expression('Risk Ratio (' *lambda[' 2ij']*'/'*lambda[' 1ij']*', Black/White)'),
       title = "Hospitalization with ADRD") +
  theme_map() +
  theme(legend.position = "bottom", 
        #legend.justification = "right", 
        #legend.direction = "vertical",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        strip.background = element_blank(),
        plot.background = element_rect(fill = "white", color = NA),
        plot.title = element_text(hjust = 0.5, size = 14)) +
  scale_fill_manual(
    # values = c("#08306b",
    #            "#2171b5",
    #            "#fee6ce",
    #            "#fdd0a2", 
    #            "#fdae6b", 
    #            "#fd8d3c", 
    #            "#f16913",
    #            "#d94801",
    #            "#a63603", 
    #            "#7f2704"),
    values = c("#08306b",
               "#313695",
               "#2171b5",
               "#74add1",
               "#abd9e9",
               "#fee6ce",
               "#fdd0a2", 
               "#fdae6b", 
               "#fd8d3c", 
               "#f16913",
               "#d94801",
               "#a63603", 
               "#7f2704"),
    na.value = "gray40",
    drop = FALSE,
    breaks = levels(unique(hosp_res_sf$risk_ratio_bins))#,
    #na.translate = F      # makes NA white
  ) +
  guides(fill = guide_legend(
    title.position = "top",
    nrow = 1,
            label.position = "bottom",
            keywidth = grid::unit(2, "lines"),
            keyheight = grid::unit(.5, "lines"),
            label.hjust = 0))

write_rds(hosp_risk_ratios_map_1m, file = "results/figures/intermediate/hosp_risk_ratios_map_1m_dual.rds")
```

```{r}
# # save plot for unadjusted model only
# # same plot but for py> 700
# 
# hosp_risk_ratios_map_1m <- hosp_res_sf %>% 
#   
#   # filter out black py < 700
#   filter(race == 2 & person_years > 700) |>
#   
#   filter(model == "hosp_m1") |>
#   select(county, geometry, model, risk_ratio_bins) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray40", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.2) +
#   labs(fill = expression('Risk Ratio (' *lambda[' 2ij']*'/'*lambda[' 1ij']*', Black/White)'),
#        title = "Hospitalization with ADRD") +
#   theme_map() +
#   theme(legend.position = "bottom", 
#         #legend.justification = "right", 
#         #legend.direction = "vertical",
#         legend.title = element_text(size = 12),
#         legend.text = element_text(size = 10),
#         strip.background = element_blank(),
#         plot.background = element_rect(fill = "white", color = NA),
#         plot.title = element_text(hjust = 0.5, size = 14)) +
#   scale_fill_manual(
#     # values = c("#08306b",
#     #            "#2171b5",
#     #            "#fee6ce",
#     #            "#fdd0a2", 
#     #            "#fdae6b", 
#     #            "#fd8d3c", 
#     #            "#f16913",
#     #            "#d94801",
#     #            "#a63603", 
#     #            "#7f2704"),
#     values = c("#08306b",
#                "#313695",
#                "#2171b5",
#                "#74add1",
#                "#abd9e9",
#                "#fee6ce",
#                "#fdd0a2", 
#                "#fdae6b", 
#                "#fd8d3c", 
#                "#f16913",
#                "#d94801",
#                "#a63603", 
#                "#7f2704"),
#     na.value = "gray40",
#     drop = FALSE,
#     breaks = levels(unique(hosp_res_sf$risk_ratio_bins))#,
#     #na.translate = F      # makes NA white
#   ) +
#   guides(fill = guide_legend(
#     title.position = "top",
#     nrow = 1,
#             label.position = "bottom",
#             keywidth = grid::unit(2, "lines"),
#             keyheight = grid::unit(.5, "lines"),
#             label.hjust = 0))
# 
# write_rds(hosp_risk_ratios_map_1m, 
#           file = "results/figures/intermediate/hosp_risk_ratios_map_1m_700py.rds")
```



```{r}
# # save plot for all three models
# 
# hosp_risk_ratios_map_3m <- hosp_res_sf %>% 
#   select(county, geometry, model, risk_ratio_bins) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray40", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.2) +
#   labs(fill = expression('Risk Ratio (' *lambda['2ij']*'/'*lambda['1ij']*', Black/White)'),
#        title = "Hospitalization with ADRD") +
#   facet_grid(~model,
#         labeller = as_labeller(model_labels)) +
#   theme_map() +
#   theme(legend.position = "bottom", 
#         #legend.justification = "right", 
#         #legend.direction = "vertical",
#         legend.title = element_text(size = 14),
#         legend.text = element_text(size = 10),
#         strip.background = element_blank(),
#         plot.title = element_text(hjust = 0.5, size = 14),
#         plot.background = element_rect(fill = "white", color = NA)) +
#   scale_fill_manual(
#     # values = c("#08306b",
#     #            "#2171b5",
#     #            "#fee6ce",
#     #            "#fdd0a2", 
#     #            "#fdae6b", 
#     #            "#fd8d3c", 
#     #            "#f16913",
#     #            "#d94801",
#     #            "#a63603", 
#     #            "#7f2704"),
#     values = c("#08306b",
#                "#313695",
#                "#2171b5",
#                "#74add1",
#                "#abd9e9",
#                "#fee6ce",
#                "#fdd0a2", 
#                "#fdae6b", 
#                "#fd8d3c", 
#                "#f16913",
#                "#d94801",
#                "#a63603", 
#                "#7f2704"),
#     na.value = "gray40",
#     drop = FALSE,
#     breaks = levels(unique(hosp_res_sf$risk_ratio_bins))#,
#     #na.translate = F      # makes NA white
#   ) +
#   guides(fill = guide_legend(
#     title.position = "top",
#     nrow = 1,
#             label.position = "bottom",
#             keywidth = grid::unit(2, "lines"),
#             keyheight = grid::unit(.5, "lines"),
#             label.hjust = 0))
# 
# write_rds(hosp_risk_ratios_map_3m, file = "results/figures/intermediate/hosp_risk_ratios_map_3m.rds")
```



```{r}
# # plot just ADRD outcome
# 
# # plot
# hosp_res_sf %>% 
#   select(county, geometry, model, risk_ratio_bins) %>%
#   distinct() %>%
#   ggplot() +
#   geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
#   geom_sf(data = county_sf_zero_b, fill = "gray40", lwd = 0.0) + 
#   geom_sf(data = state_sf, fill = NA, col = "black", lwd = 0.1) +
#   labs(fill = "Risk Ratio \n(Black/White)") +
#   facet_grid(~ model,
#              labeller = as_labeller(model_labels)
#              ) +
#   theme_map() +
#   theme(legend.position = "bottom", 
#         #legend.justification = "right", 
#         #legend.direction = "vertical",
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 8),
#         strip.background = element_blank()) +
#   scale_fill_manual(
#     # values = c("#08306b",
#     #            "#2171b5",
#     #            "#fee6ce",
#     #            "#fdd0a2", 
#     #            "#fdae6b", 
#     #            "#fd8d3c", 
#     #            "#f16913",
#     #            "#d94801",
#     #            "#a63603", 
#     #            "#7f2704"),
#     values = c("#08306b",
#                "#313695",
#                "#2171b5",
#                "#74add1",
#                "#abd9e9",
#                "#fee6ce",
#                "#fdd0a2", 
#                "#fdae6b", 
#                "#fd8d3c", 
#                "#f16913",
#                "#d94801",
#                "#a63603", 
#                "#7f2704"),
#     na.value = "gray40",
#     drop = FALSE,
#     breaks = levels(unique(hosp_res_sf$risk_ratio_bins))#,
#     #na.translate = F      # makes NA white
#   ) +
#   guides(fill = guide_legend(
#     title.position = "top",
#     nrow = 1,
#             label.position = "bottom",
#             keywidth = grid::unit(2.5, "lines"),
#             keyheight = grid::unit(.5, "lines"),
#             label.hjust = 0))
# 
# # # or change legend to be along the bottom like Matt's
# # hosp_res_sf %>% 
# #   select(county, geometry, risk_ratio_bins) %>%
# #   distinct() %>%
# #   ggplot() +
# #   geom_sf(aes(fill = risk_ratio_bins), lwd = 0.03) + 
# #   #lakes_gg +
# #   states_gg +
# #   geom_sf(data = lakes_sf, fill = "white", col = "black") +
# #   labs(fill = "Risk Ratio (Black/White)") +
# #   theme_map() +
# #   theme(legend.position = "bottom", 
# #         #legend.justification = "right", 
# #         #legend.direction = "vertical",
# #         legend.title = element_text(size = 10),
# #         legend.text = element_text(size = 8)) +
# #   scale_fill_manual(values = c("#07da63", 
# #                                "#7ffbb5",
# #                                "#FDE0EF", 
# #                                "#fbc8e3", 
# #                                #"#fab0d6", 
# #                                "#DE77AE", 
# #                                "#C51B7D", 
# #                                "#8E0152", 
# #                                "#400023"),
# #                     na.value = "gray50",
# #                     breaks = levels(unique(hosp_res_sf$risk_ratio_bins))) +
# #   guides(fill = guide_legend(
# #     title.position = "top",        
# #     nrow = 1,
# #             label.position = "bottom",
# #             keywidth = grid::unit(3.5, "lines"),
# #             keyheight = grid::unit(.5, "lines"),
# #             label.hjust = 0))

```


## Look at distributions of post-modeling variables across counties

### In a table

group by race doesn't actually affect grouping


### DELETED EVERYTHING BELOW THIS
