
# Explore num enrollees

```{r, warning=F, message=F}
## The interactive Rstudio session was started with 38 GB and 8 cores
## The  default working directory of an Rmd in Rstudio is the Rmd location

## load packages ----
library(tidyverse)
library(magrittr)
library(sf)
```

### Read shp files

* read county shp files

```{r}
county_sf <- read_sf("../data/input/scratch/tl_2022_us_county/tl_2022_us_county.shp")

dim(county_sf)
names(county_sf)
```


* read zcta shp files

```{r}
zcta_sf <- read_sf("../data/input/scratch/tl_2022_us_zcta520/tl_2022_us_zcta520.shp")

dim(zcta_sf)
names(zcta_sf)
```

### Read zip to zcta crosswalk file

```{r}
zctazipcrosswalk <- read_csv("../data/input/zip_to_zcta/Zip_to_ZCTA_crosswalk_2015_JSI.csv")
```

```{r}
dim(zctazipcrosswalk)
length(unique(zctazipcrosswalk$ZIP))
length(unique(zctazipcrosswalk$ZCTA))
```

* Obtain crosswalk weights

```{r}
zctazipcrosswalk %<>% 
  left_join(
    zctazipcrosswalk %>% 
      select(ZIP, ZCTA) %>% 
      group_by(ZCTA) %>% 
      summarise(n = n()) %>% 
      mutate(w = 1/n)
  )

head(zctazipcrosswalk)
```

### Read num enrollees

```{r}
fips_num_enroll <- read_csv("../data/input/fips_num_enroll.csv")
zips_num_enroll <- read_csv("../data/input/zips_num_enroll.csv")
```

```{r}
dim(fips_num_enroll)
dim(zips_num_enroll)
```

### Prepare datasets

* corroborate both zips and zip counts consider the same number of enrollees

```{r}
sum(fips_num_enroll$count)
sum(zips_num_enroll$count)
```

* mapping fips in the shp files

Some fips not found in shp files.

```{r}
length(unique(fips_num_enroll$fips5))
sum(!unique(fips_num_enroll$fips5) %in% county_sf$GEOID)
```

* mapping zips into zctas into shp files

Some zip codes not found in crosswalk file.

```{r}
length(unique(as.numeric(zips_num_enroll$zip)))
sum(!unique(as.numeric(zips_num_enroll$zip)) %in% zctazipcrosswalk$ZIP)
```

There are zctas not found in shp files

```{r}
zips_num_enroll %<>% 
  left_join(
    zctazipcrosswalk %>% 
      mutate(ZIP = as.numeric(ZIP)), 
    by = c("zip" = "ZIP")
  )

length(unique(zips_num_enroll$ZCTA))
sum(!unique(zips_num_enroll$ZCTA) %in% zcta_sf$ZCTA5CE20)
```

### Explore race

* corroborate both zips and zip race counts give the same numbers

```{r}
tapply(fips_num_enroll$count, fips_num_enroll$race, sum)
```

```{r}
tapply(fips_num_enroll$count, fips_num_enroll$race, sum)
```

* look at proportions by race

77% white and 19% black

```{r}
prettyNum(prop.table(tapply(fips_num_enroll$count, fips_num_enroll$race, sum)))
```

### Explore num enrollees

* All counties have enrollee counts in all years

```{r}
xx <- fips_num_enroll %>% 
  group_by(year, fips5) %>% 
  summarise(count = sum(count)) %>% 
  group_by(fips5) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  )

summary(xx$n)
```

* Pct change through years by county

```{r}
xx %>% 
  arrange(desc(pct_diff))
```

* Summary of mean enrollees per county

```{r}
summary(xx$mean) 
```

* Map num enrollees per county

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    xx %>%  
      mutate(fips5 = as.character(fips5)), 
    by = c("GEOID" = "fips5")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = mean), aes = 0.5)
```

* Zip's have varying number of years with enrollees

Zip with zero enrollees are not included.

```{r}
xx <- zips_num_enroll %>% 
  group_by(year, zip) %>% 
  summarise(count = sum(count)) %>% 
  group_by(zip) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  ) 

summary(xx$n) 
```

* Pct change through years by zip

```{r}
xx %>% 
  arrange(desc(pct_diff))
```

* Summary of mean enrollees per zip

To obtain the mean through year, zip-year having enrollee counts of zero have to be included.

```{r}
summary(xx$mean) 
```

* Map num enrollees per zip

Zips to Zcta crosswalk or shapefiles seem to be outdated.

```{r}
xx <- zips_num_enroll %>% 
  group_by(year, ZCTA) %>% 
  summarise(count = sum(count * w)) %>% 
  group_by(ZCTA) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  ) 

zcta_sf %>% 
  select(ZCTA5CE20) %>% 
  right_join(
    xx, 
    by = c("ZCTA5CE20" = "ZCTA")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = mean), aes = 0.5)
```

### Enrollee pop per county per race

* County-race (white & black) have varying number of years with enrollees. County-race-year with zero counts are excluded

```{r}
xx <- fips_num_enroll %>% 
  filter(race %in% c(1,2)) %>% 
  group_by(year, fips5, race) %>% 
  summarise(count = sum(count)) %>% 
  ungroup() %>% 
  group_by(fips5, race) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  ) %>% 
  ungroup()

tapply(xx$n, xx$race, summary)
```

* Pct change through years by county

```{r}
xx %>% 
  arrange(desc(pct_diff))
```

* Summary of mean enrollees per county-race

```{r}
tapply(xx$mean, xx$race, summary)
```

* Map num enrollees per county

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    xx %>%  
      mutate(fips5 = as.character(fips5)), 
    by = c("GEOID" = "fips5")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = mean), aes = 0.5) +
  facet_wrap(~race)
```
