# ADRD hospitalization per county

## Limited to NC

```{r, message=F, warning=F}
## The interactive Rstudio session was started with 36 GB and 8 cores
## The  default working directory of an Rmd in Rstudio is the Rmd location
## The details of the symbolic links to the input data are provided in the README of the  /data/input folder

## load packages ----
library(tidyverse)
library(magrittr)
library(fst)
library(data.table)
library(sf)
library(tictoc)
```

### Read num enrollees

```{r}
fips_num_enroll_ <- read_csv("../data/input/fips_num_enroll_.csv")
```

### Read shp files

* read county shp files

```{r}
county_sf <- read_sf("../data/input/scratch/tl_2022_us_county/tl_2022_us_county.shp")

dim(county_sf)
names(county_sf)
```

### Read SSA5 to FIPS

```{r}
ssa_fips <- read_csv("../data/input/ssa_fips_state_county2016.csv")
```

```{r}
dim(ssa_fips)
length(unique(ssa_fips$ssacounty))
length(unique(ssa_fips$fipscounty))
```

### Read ADRD admissions

```{r}
tic("Read ADRD admissions")
admissions <- read_csv("../data/intermediate/admissions.csv")
toc()
```

```{r}
class(admissions)
dim(admissions)
lapply(admissions, class)
```

* crosswalk ssa5 and fips

some counties not in crosswalk file

```{r}
sum(!unique(as.numeric(admissions$SSA5)) %in% ssa_fips$ssacounty)
```

```{r}
admissions %<>% 
  mutate(SSA5 = as.character(SSA5)) %>% 
  left_join(
    select(ssa_fips, ssacounty, fipscounty), 
    by = c("SSA5" = "ssacounty"))
```

### Prepare data

* Mapping of enrollment fips with admissions fips

All admission fips in enrollments fips

```{r}
sum(!unique(as.numeric(admissions$fipscounty)) %in% fips_num_enroll_$fips5)
```

### Explore race

* fips race counts

```{r}
tapply(admissions$counts, admissions$RACE, sum)
```

* look at proportions by race

76% white and 21% black

```{r}
prettyNum(prop.table(tapply(admissions$counts, admissions$RACE, sum)))
```

### Explore num hospitalizations

* All counties have adrd counts in all years. 

Crosswalk may be required by year

```{r}
xx <- admissions %>% 
  group_by(year, SSA5) %>% 
  summarise(count = sum(counts)) %>% 
  group_by(SSA5) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  )

summary(xx$n)
```

```{r}
xx <- admissions %>% 
  group_by(year, fipscounty) %>% 
  summarise(count = sum(counts)) %>% 
  group_by(fipscounty) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  )

summary(xx$n)
```

* Pct change through years by county

```{r}
xx %>% 
  arrange(desc(pct_diff))
```

* Summary of mean hospitalizations per county

```{r}
summary(xx$mean) 
```

* Map num hospitalizations per county

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    xx, 
    by = c("GEOID" = "fipscounty")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = mean), aes = 0.5)
```

### Hospitalizations per county per race

* County-race (white & black) have varying number of years with hospitalizations. County-race-year with zero counts are excluded

```{r}
xx <- admissions %>% 
  filter(RACE %in% c(1,2)) %>% 
  group_by(year, fipscounty, RACE) %>% 
  summarise(count = sum(counts)) %>% 
  ungroup() %>% 
  group_by(fipscounty, RACE) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    mean = mean(count, na.rm = T), 
    pct_diff = max / min - 1 
  ) %>% 
  ungroup()

tapply(xx$n, xx$RACE, summary)
```

* Pct change through years by county

```{r}
xx %>% 
  arrange(desc(pct_diff))
```

* Summary of hospitalizations per county-race

```{r}
tapply(xx$mean, xx$RACE, summary)
```

* Map num enrollees per county

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  right_join(
    xx %>%  
      mutate(fips5 = as.character(fipscounty)), 
    by = c("GEOID" = "fipscounty")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = mean), aes = 0.5) +
  facet_wrap(~RACE)
```
