# ADRD hospitalization rates per county

## Limited to NC

```{r, message=F, warning=F}
## The interactive Rstudio session was started with 36 GB and 8 cores
## The  default working directory of an Rmd in Rstudio is the Rmd location
## The details of the symbolic links to the input data are provided in the README of the  /data/input folder

## load packages ----
library(tidyverse)
library(magrittr)
library(fst)
library(data.table)
library(sf)
library(tictoc)
```

### Read shp files

* read county shp files

```{r}
county_sf <- read_sf("../data/input/scratch/tl_2022_us_county/tl_2022_us_county.shp")

dim(county_sf)
names(county_sf)
```

### Read zcta to fips files

```{r}
countyzctacrosswalk <- read_tsv("../data/input/countyzctacrosswalk.tab")
```

* Obtain weights

```{r}
xx <- countyzctacrosswalk %>%
  select(county, ZCTA5) %>%
  group_by(ZCTA5) %>%
  summarise(n = n()) %>%
  mutate(w = 1/n)

countyzctacrosswalk %<>%
  left_join(xx)

head(countyzctacrosswalk)
```

### Read zip to zcta

```{r}
zipzctacrosswalk <- read_csv("../data/input/zip_to_zcta/Zip_to_ZCTA_crosswalk_2015_JSI.csv") %>% 
  rename(
    zcta = ZCTA,
    zip = ZIP
  )
```

* check mapping properties

```{r}
dim(zipzctacrosswalk)
length(unique(zipzctacrosswalk$zip))
length(unique(zipzctacrosswalk$zcta))
```

```{r}
xx <- zipzctacrosswalk %>%
  select(zip, zcta) %>%
  group_by(zcta) %>%
  summarise(n = n()) %>%
  mutate(w = 1/n)

zipzctacrosswalk %<>%
  left_join(xx)

head(zipzctacrosswalk)
```

### Read ADRD events

```{r}
tic("Read denom files")
ffs_entry_exit_adrd <- read_fst("../data/input/denom/ffs_entry_exit_adrd.fst" 
                                #, as.data.table = TRUE
)
toc()
```

```{r}
class(ffs_entry_exit_adrd)
dim(ffs_entry_exit_adrd)
lapply(ffs_entry_exit_adrd, class)
```

* determine unique ID's.

A single row per ID

```{r}
tic("sum duplicates")
sum(duplicated(ffs_entry_exit_adrd$qid))
toc()
```

* Filter beneficiaries with ADRD

```{r}
adrd_df <- ffs_entry_exit_adrd %>% 
  filter(ADRD_hosp) %>% 
  select(qid, race_cat, sex, ADRD_year, ADRD_zip, ADRD_age)

dim(adrd_df)
```

* explore race

80% white, 18% black

```{r}
prop.table(table(adrd_df$race_cat))
```

### Prepare data

* Mapping of ADRD_zip with zcta and state and county

Some zips not found in crosswalk to zcta (all US)

```{r}
length(unique(adrd_df$ADRD_zip))
sum(!unique(as.character(adrd_df$ADRD_zip)) %in% zipzctacrosswalk$zip)
```

Some zctas codes not found in crosswalk to county (all US)

```{r}
length(unique(zipzctacrosswalk$zcta))
sum(!unique(zipzctacrosswalk$zcta) %in% countyzctacrosswalk$ZCTA5)
```

Some zero zip codes

```{r}
sum(as.numeric(adrd_df$ADRD_zip) == 0, na.rm = T)
```

The left join produces duplicated q_id entries

Beneficiaries that live in zips that intersect more than one county contribute more than once. The weights distribute them.

```{r}
tic("map ADRD_zip")
adrd_df %<>%
  mutate(ADRD_zip = as.character(ADRD_zip)) %>%
  left_join(
    zipzctacrosswalk,
    by = c("ADRD_zip" = "zip")
  )
toc()
```

```{r}
tic("map ADRD_zip")
adrd_df %<>%
  left_join(
    countyzctacrosswalk %>% 
      mutate(ZCTA5 = as.character(ZCTA5)),
    by = c("zcta" = "ZCTA5")
  )
toc()
```

* Limit to NC

```{r}
adrd_df %<>% 
  filter(state == "NC")

dim(adrd_df)
```

* harmonize weights

```{r}
adrd_df %<>% 
  mutate(w = w.x * w.y)
```


### ADRD historic first hospitalizations in NC

* Summarize number of historic first ADRD hospitalizations per county

```{r}
yy <- adrd_df %>% 
  group_by(FIPS) %>% 
  summarize(adrd_count = sum(w))

summary(yy$adrd_count)
```

* Map adrd counts

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    yy %>%  
      mutate(FIPS = as.character(FIPS)), 
    by = c("GEOID" = "FIPS")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = adrd_count), aes = 0.5)
```

### ADRD historic first hospitalizations per race

* Summarize number of historic first ADRD hospitalizations per county-race

```{r}
yy <- adrd_df %>% 
  group_by(FIPS, race_cat) %>% 
  summarize(adrd_count = sum(w)) %>% 
  ungroup()

tapply(yy$adrd_count, yy$race_cat, summary)
```

* Map adrd counts

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    yy %>%  
      filter(race_cat %in% c("white", "black")) %>% 
      mutate(FIPS = as.character(FIPS)), 
    by = c("GEOID" = "FIPS")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = adrd_count)) +
  facet_grid(~ race_cat)
```

```{r}
write_rds(adrd_df, "../data/intermediate/adrd_df.rds")
```

