# ADRD hospitalization rates per county

## Limited to NC

```{r, message=F, warning=F}
## The interactive Rstudio session was started with 36 GB and 8 cores
## The  default working directory of an Rmd in Rstudio is the Rmd location
## The details of the symbolic links to the input data are provided in the README of the  /data/input folder

## load packages ----
library(tidyverse)
library(magrittr)
library(fst)
library(data.table)
library(sf)
library(tictoc)
```

### Read num enrollees

```{r}
fips_num_enroll_ <- read_csv("../data/input/fips_num_enroll_.csv") %>% 
  rename(fips = fips5, 
         num_enroll = count) %>% 
  mutate(fips = as.character(fips))
```

### Read shp files

* read county shp files

```{r}
county_sf <- read_sf("../data/input/scratch/tl_2022_us_county/tl_2022_us_county.shp") %>% 
  rename(fips = GEOID) %>% 
  mutate(fips = as.character(as.numeric(fips)))

dim(county_sf)
names(county_sf)
```

### Read SSA5 to FIPS

```{r}
ssa_fips <- read_csv("../data/input/ssa_fips_state_county2016.csv") %>% 
  rename(ssa = ssacounty, 
         fips = fipscounty)
```

```{r}
dim(ssa_fips)
length(unique(ssa_fips$ssa))
length(unique(ssa_fips$fips))
```

### Read ADRD admissions

```{r}
admissions <- read_csv("../data/intermediate/admissions.csv") %>% 
  rename(ssa = SSA5, 
         race = RACE, 
         adrd_hosp = counts) %>% 
  mutate(ssa = as.character(ssa))
```

```{r}
class(admissions)
dim(admissions)
```

* crosswalk ssa5 and fips

some counties not in crosswalk file

```{r}
sum(!unique(as.numeric(admissions$ssa)) %in% ssa_fips$ssa)
```

```{r}
admissions %<>% 
  left_join(
    select(ssa_fips, ssa, fips))
```

### Prepare data

* Mapping of enrollment fips with admissions fips

```{r}
sum(!unique(admissions$fips) %in% fips_num_enroll_$fips)
sum(!unique(fips_num_enroll_$fips) %in% admissions$fips )
```

```{r}
fips_rates <- fips_num_enroll_ %>% 
  left_join(admissions) %>% 
  mutate(rates = adrd_hosp / num_enroll)
```

* there are counties with quite a large enrollee populations that are not mapped to hospitalizations using (but maybe they do with correct ssa?).

```{r}
fips_rates %>% 
  filter(is.na(rates)) %>% 
  select(fips, race, year, num_enroll) %>% 
  pivot_wider(
    id_cols = c('fips', 'race'), 
    values_from = 'num_enroll', 
    names_from = 'year') %>% 
  DT::datatable()
```

### Explore rates per race

* summary

```{r}
tapply(fips_rates$rates, fips_rates$race, summary)
```

* some fips do not have rates for all years

```{r}
xx <- fips_rates %>%  
  filter(
    !is.na(rates), 
    !is.na(fips)
  ) %>% 
  group_by(fips, race) %>% 
  summarise(
    n = n(),
    max = max(rates, na.rm = T), 
    min = min(rates, na.rm = T), 
    mean = mean(rates, na.rm = T), 
    pct_diff = max / min - 1 
  )

summary(xx$n)
```

* Pct change through years by county

```{r}
summary(xx$pct_diff)
```

* County 37990 is suspicious.

```{r}
fips_rates %>% 
  filter(
    !is.na(rates), 
    !is.na(fips)
  ) %>% 
  left_join(xx) %>% 
  arrange(desc(pct_diff, year))
```

* time series of a sample of hospitalizations per county

```{r}
yy <- fips_rates %>% 
  filter(
    !fips == '37990', 
    !is.na(rates), 
    !is.na(fips)
  ) %>% 
  left_join(xx) %>% 
  arrange(desc(pct_diff, year)) 

yy[1:170,] %>% 
  ggplot() + 
  geom_line(aes(x = year, y = rates, color = fips, linetype = as.factor(race)))
```

* Map mean rates per county

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(fips) %>% 
  left_join(xx) %>% 
  filter(!is.na(race)) %>% 
  ggplot() + 
  geom_sf(aes(fill = mean), aes = 0.5) + 
  facet_grid(~race)
```
