
# Explore num enrollees

```{r, warning=F, message=F}
## The interactive Rstudio session was started with 38 GB and 8 cores
## The  default working directory of an Rmd in Rstudio is the Rmd location

## load packages ----
library(tidyverse)
library(magrittr)
library(sf)
```

### Read shp files

* read county shp files

```{r}
county_sf <- read_sf("../data/input/scratch/tl_2022_us_county/tl_2022_us_county.shp")

dim(county_sf)
names(county_sf)
```


* read zcta shp files

```{r}
zcta_sf <- read_sf("../data/input/scratch/tl_2022_us_zcta520/tl_2022_us_zcta520.shp")

dim(zcta_sf)
names(zcta_sf)
```

### Read num enrollees

```{r}
fips_num_enroll <- read_csv("../data/input/fips_num_enroll.csv")
zcta_num_enroll <- read_csv("../data/input/zcta_num_enroll.csv")
```

```{r}
dim(fips_num_enroll)
dim(zcta_num_enroll)
```

* mapping fips in the shp files

There are fips not found in shp files

```{r}
length(unique(fips_num_enroll$fips5))
sum(!unique(fips_num_enroll$fips5) %in% county_sf$GEOID)
```

* mapping zctas are in the shp files

There are zctas not found in shp files

```{r}
length(unique(zcta_num_enroll$zip))
sum(unique(zcta_num_enroll$zip) %in% zcta_sf$ZCTA5CE20)
```

### Explore race

```{r}
table(fips_num_enroll$race)
```

```{r}
prop.table(table(fips_num_enroll$race))
```

### Explore num enrollees

* All counties have enrollee counts in all years

```{r}
xx <- fips_num_enroll %>% 
  group_by(year, fips5) %>% 
  summarise(count = sum(count)) %>% 
  group_by(fips5) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    pct_diff = max / min - 1 
  )

summary(xx$n)
```

* Pct change through years by county

```{r}
xx %>% 
  arrange(desc(pct_diff))
```


* Zcta's have varying number of years with enrollees

To obtain the mean through year, zcta-year having enrollee counts of zero have to be included.

```{r}
xx <- zcta_num_enroll %>% 
  group_by(year, zip) %>% 
  summarise(count = sum(count)) %>% 
  group_by(zip) %>% 
  summarise(
    n = n(),
    max = max(count, na.rm = T), 
    min = min(count, na.rm = T), 
    pct_diff = max / min - 1 
  ) 

summary(xx$n) 
```

```{r}
xx %>% 
  arrange(desc(pct_diff))
```

### Enrollee pop per county

```{r}
xx <- fips_num_enroll %>% 
  group_by(year, fips5) %>% 
  summarise(count = sum(count)) %>% 
  group_by(fips5) %>% 
  summarise(
    num_enrollee = mean(count, na.rm = T) 
  ) 

summary(xx$num_enrollee)
```

```{r}
hist(xx$num_enrollee)
```

* Map num enrollees

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    xx %>%  
      mutate(fips5 = as.character(fips5)), 
    by = c("GEOID" = "fips5")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = num_enrollee), aes = 0.5)
```

### Enrollee pop per county per race

```{r}
# xx <- fips_num_enroll %>% 
#   group_by(year, fips5, race) %>% 
#   summarise(count = sum(count)) %>% 
#   group_by(fips5, race) %>% 
#   summarise(
#     num_enrollee = mean(count, na.rm = T) 
#   ) %>%
#   group_by(race)
#   nest()
# 
# lapply(xx$num_enrollee, summary)
```







