# ADRD hospitalization rates per county

## Limited to NC

```{r, message=F, warning=F}
## The interactive Rstudio session was started with 36 GB and 8 cores
## The  default working directory of an Rmd in Rstudio is the Rmd location
## The details of the symbolic links to the input data are provided in the README of the  /data/input folder

## load packages ----
library(tidyverse)
library(magrittr)
library(fst, lib.loc = ".")
library(data.table)
library(sf)
library(tictoc, lib.loc = ".")
```

### Read num enrollees

```{r}
fips_num_enroll <- read_csv("../data/input/fips_num_enroll.csv")
zcta_num_enroll <- read_csv("../data/input/zcta_num_enroll.csv")
```

### Read shp files

* read county shp files

```{r}
county_sf <- read_sf("../data/input/scratch/tl_2022_us_county/tl_2022_us_county.shp")

dim(county_sf)
names(county_sf)
```

* read zcta shp files

```{r}
zcta_sf <- read_sf("../data/input/scratch/tl_2022_us_zcta520/tl_2022_us_zcta520.shp")

dim(zcta_sf)
names(zcta_sf)
```

### Read zcta to fips crosswalk file

```{r}
countyzipcrosswalk <- read_tsv("../data/input/countyzipcrosswalk.tab")
```

* Obtain weights

```{r}
xx <- countyzipcrosswalk %>% 
  select(county, ZCTA5) %>% 
  group_by(ZCTA5) %>% 
  summarise(n = n()) %>% 
  mutate(w = 1/n)

countyzipcrosswalk %<>% 
  left_join(xx)

head(countyzipcrosswalk)
```

### Read ADRD events

```{r}
tic("Read denom files")
ffs_entry_exit_adrd <- read_fst("../data/input/denom/ffs_entry_exit_adrd.fst" 
                                #, as.data.table = TRUE
)
toc()
```

```{r}
class(ffs_entry_exit_adrd)
dim(ffs_entry_exit_adrd)
lapply(ffs_entry_exit_adrd, class)
```

* determine unique ID's.

A single row per ID

```{r}
tic("sum duplicates")
sum(duplicated(ffs_entry_exit_adrd$qid))
toc()
```

### Prepare data

* Filter beneficiaries with ADRD

```{r}
adrd_df <- ffs_entry_exit_adrd %>% 
  filter(ADRD_hosp) %>% 
  select(qid, race_cat, sex, ADRD_year, ADRD_zip, ADRD_age)

dim(adrd_df)
```

* Mapping of ADRD_zip with state and county

Some zip codes not found in crosswalk file

```{r}
sum(!unique(as.numeric(adrd_df$ADRD_zip)) %in% countyzipcrosswalk$ZCTA5)
```

Some zero zip codes

```{r}
sum(as.numeric(adrd_df$ADRD_zip) == 0, na.rm = T)
```

The left join produces duplicated q_id entries

Beneficiaries that live in zips that intersect more than one county contribute more than once. The weights distribure them.

```{r}
tic("map ADRD_zip")
adrd_df %<>%
  mutate(ADRD_zip = as.numeric(ADRD_zip)) %>%
  left_join(
    countyzipcrosswalk,
    by = c("ADRD_zip" = "ZCTA5")
  )
toc()
```

* Limit to NC

```{r}
adrd_df %<>% 
  filter(state == "NC")
```

### ADRD hospitalizations in NC

* Summarize number of ADRD cases per county

```{r}
yy <- adrd_df %>% 
  group_by(FIPS) %>% 
  summarize(adrd_count = sum(w))

summary(yy$adrd_count)
```

* Map adrd counts

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    yy %>%  
      mutate(FIPS = as.character(FIPS)), 
    by = c("GEOID" = "FIPS")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = adrd_count), aes = 0.5)
```

### Compute ADRD rates per county

```{r}
xx <- fips_num_enroll %>% 
  group_by(year, fips5) %>% 
  summarise(count = sum(count)) %>% 
  group_by(fips5) %>% 
  summarise(
    num_enrollee = mean(count, na.rm = T) 
  ) 

zz <- left_join(xx, yy, by = c("fips5" = "FIPS")) %>% 
  mutate(adrd_rate = adrd_count / num_enrollee)

summary(zz$adrd_rate)
```

* Map adrd rates

```{r}
county_sf %>% 
  filter(STATEFP == "37") %>% 
  select(GEOID) %>% 
  left_join(
    zz %>%  
      mutate(fips5 = as.character(fips5)), 
    by = c("GEOID" = "fips5")
  ) %>% 
  ggplot() + 
  geom_sf(aes(fill = sqrt(adrd_rate)))
```
